---
title: "Assignment 4"
author: "Chul Hee Kim"
date: '2021 10 2 '
output: pdf_document
---

```{r include=FALSE, warning=FALSE, message=FALSE}
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(MASS)) install.packages("MASS", repos = "http://cran.us.r-project.org")
if(!require(onewaytests)) install.packages("onewaytests", repos = "http://cran.us.r-project.org")
if(!require(faraway)) install.packages("faraway", repos = "http://cran.us.r-project.org")
if(!require(knitr)) install.packages("faraway", repos = "http://cran.us.r-project.org")
if(!require(kableExtra)) install.packages("kableExtra", repos = "http://cran.us.r-project.org")
library(tidyverse)
library(caret)
library(MASS)
library(onewaytests)
library(faraway)
library(knitr)
library(kableExtra)
options(scipen=999)
```

## Problem 1

a-)
```{r P1_a}
CDI.data <- read.csv("/cloud/project/Data/CDI Data.csv")
Y <- CDI.data$Number.of.active.physicians
X1 <- CDI.data$Total.population
X2 <- CDI.data$Total.personal.income
X3 <- I(CDI.data$Geographic.region == 1)*1 # NE
X4 <- I(CDI.data$Geographic.region == 2)*1 # NC
X5 <- I(CDI.data$Geographic.region == 3)*1 # S
CDI.data2 <- data.frame(cbind(Y,X1,X2,X3,X4,X5))
CDI.reg <- lm(Y ~ ., data=CDI.data2)
summary(CDI.reg)

n <- nrow(CDI.data2) # 440
p <- 6
b3 <- CDI.reg$coefficients[4] # 149.02
b4 <- CDI.reg$coefficients[5] # 145.53
T.crit <- qt(0.975,n-p) # 1.965
b3.se <- summary(CDI.reg)$coefficients[4,2] # 86.833
b4.se <- summary(CDI.reg)$coefficients[5,2] # 85.153
b3.lwr <- b3 - T.crit*b3.se # 5.89
b3.upr <- b3 + T.crit*b3.se # 292.15
b4.lwr <- b4 - T.crit*b4.se # 5.16
b4.upr <- b4 + T.crit*b4.se # 285.89

# Confirm the results with 'confint' function 
confint(CDI.reg)[4:5,]
```
Confidence Intervals (calculated manually): \newline
`r round(b3.lwr,3)` $\le \beta_3 \le$ `r round(b3.upr,3)` \newline
`r round(b4.lwr,3)` $\le \beta_4 \le$ `r round(b4.upr,3)`

The confidence intervals for $\beta_3$ and $\beta4$ are very similar and mostly 
overlap with each other, so the effect for NE region on number of active physicians 
does not differ from the effect for NC region.

$*$ Below, I attempt to create the confidence interval for $\beta_4-\beta_3$
```{r P1_a2}
b4b3 <- b4 - b3 # -3.493
b4.b3.se <- vcov(CDI.reg)[5,4] # 4289.97
b4.var <- vcov(CDI.reg)[5,5] # 7251.02
b3.var <- vcov(CDI.reg)[4,4] # 7540
b4b3.var <- b4.var + b3.var - 2*b4.b3.se # 6211.09
b4b3.lwr <- b4b3 - T.crit*sqrt(b4b3.var) # -158.39
b4b3.upr <- b4b3 + T.crit*sqrt(b4b3.var) # 151.40
```
$(b_4-b_3) \pm t(1-\frac{\alpha}{2},df)s\{b_4-b_3\}$ \newline
$s^2\{b_4-b_3\} = s^2\{b_4\} + s^2\{b_3\} - 2s\{b_4,b_3\}$

Confidence Interval:
`r round(b4b3.lwr,3)` $\le \beta_4 - \beta_3 \le$ `r round(b4b3.upr,3)`

The confidence interval above includes zero, so we cannot conclude that the effect
for the NE region on number of physicians differs from that for NC region.
It is consistent with the result from comparing two confidence intervals above.

b-)
```{r P1_b}
# Reduced model
CDI.reg.reduced <- lm(Y ~ X1+X2, data=CDI.data2)
SSE_F <- anova(CDI.reg)$`Sum Sq`[6]
SSE_R <- anova(CDI.reg.reduced)$`Sum Sq`[3]
dfF <- 434 # n - 6
dfR <- 437 # n - 3
F.stat <- ((SSE_R-SSE_F)/(dfR-dfF)) / (SSE_F/dfF) # 1.949
F.crit <- qf(0.9,dfR-dfF,dfF) # 2.096
pv <- 1-pf(F.stat,3,434) # 0.121
```
Full Model: $Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + \beta_4X_4 + \beta_5X_5$ \newline
Reduced Model: $Y = \beta_0 + \beta_1X_1 + \beta_2X_2$

$H_0: \beta_3 = \beta_4 = \beta_5 = 0$ \newline
$H_a$: not all $\beta_3$ and $\beta_4$ and $\beta_5$ equal zero

$F^* = \frac{SSE(R)-SSE(F)}{df_R-df_F}\div\frac{SSE(F)}{df_F} =$ `r round(F.stat,3)` \newline
$F(1-\alpha;df_R-df_F,df_f)$ = F(0.95;3,434) = `r round(F.crit,3)`

$F^* \le F(0.95;3,434)$, so we conclude $H_0$ that no geographic effects are present.
p-value = `r round(pv,3)`.

Below is the simpler method to get the $F^*$ and p-value.
```{r P1_b2}
anova(CDI.reg, CDI.reg.reduced)
```
$F^* = 1.949$ and p-value = 0.121 are the same as the one we calculated manually.


## Problem 2

a-)
```{r P2_a}
GR.data <- read.csv("/cloud/project/Data/Grocery Retailer Data Set.csv")
GR.reg <- lm(Y ~ X1+X3+X2, data=GR.data)
summary(GR.reg)
b0 <- GR.reg$coefficients[1] # 4,149.89
b1 <- GR.reg$coefficients[2] # 0.000787
b2 <- GR.reg$coefficients[4] # -13.166
b3 <- GR.reg$coefficients[3] # 623.55
```
The estimated regression function: 
Y = `r round(b0,3)` + `r round(b1,3)`$X_1$ + `r round(b2,3)`$X_2$ + `r round(b3,3)`$X_3$

The interpretation of the coefficients: \newline
$b_1$: the change in E[Y] with a unit change in $X_1$ while $X_2$ and $X_3$ are held constant \newline
$b_2$: the change in E[Y] with a unit change in $X_2$ while $X_1$ and $X_3$ are held constant \newline
$b_3$: the change in E[Y] with a unit change in $X_3$ while $X_1$ and $X_2$ are held constant
\newpage
b-)
```{r P2_b}
plot(GR.reg$residuals, xlab="Time (weeks)", ylab="Residuals")
```

The time plot of residuals shown above seems to have no pattern and shows no indication
that the error terms are correlated.

c-)
```{r P2_c}
anova(GR.reg)
SSR.X1 <- anova(GR.reg)$`Sum Sq`[1] # 136,366
SSR.X3_1 <- anova(GR.reg)$`Sum Sq`[2] # 2,033,565
SSR.X2_13 <- anova(GR.reg)$`Sum Sq`[3] # 6,675
```
$SSR(X_1)$ = `r round(SSR.X1,0)` \newline
$SSR(X_3\vert X_1)$ = `r round(SSR.X3_1,0)` \newline
$SSR(X_2\vert X_1,X_3)$ = `r round(SSR.X2_13,0)`

d-)
```{r P2_d}
GR.reg.reduced <- lm(Y ~ X1+X3, data=GR.data)
anova(GR.reg, GR.reg.reduced)
F1 <- anova(GR.reg, GR.reg.reduced)[2,5] # F = 0.3251
F.crit <- qf(0.95,1,48) # 4.04
pvalue <- 1 - pf(F1,1,48) # 0.5712
```
$H_0: \beta_2 = 0$
$H_a: \beta_2 \ne 0$

ANOVA table shows that $F^*$ = 0.3251 $\le$ F(0.95;1,48) = `r round(F.crit,2)`,
so we conclude $H_0$ that $X_2$ can be dropped from the regression model that
already contains $X_1$ and $X_3$. The p-value is `r round(pvalue,3)` and is greater than the
significant level ($\alpha$=0.05).

$*$Alternative way to get $F^*$
```{r}
F2 <- anova(GR.reg)$`Mean Sq`[3] / anova(GR.reg)$`Mean Sq`[4]
```
$F^* = \frac{MSR(X_2\vert X_1,X_3)}{MSE(X_1,X_2,X_3)}$ (p.264 of textbook) \newline
$F^*$ = `r round(F2,4)`


## Problem 3

```{r P3_1}
data(ozone)
ozone.reg <- lm(O3 ~ temp+humidity+ibh, data=ozone)

par(mfrow=c(1,2))
plot(ozone.reg$fitted.values, ozone.reg$residuals, xlab="Fitted Values", ylab="Residuals")
error.std3 <- rstandard(ozone.reg)
qqnorm(error.std3, xlab="Normal Scores", ylab="Standardized Residuals", 
       main="Normal Probability Plot")
qqline(error.std3, col="steelblue", lwd=2)
```

QQ plot shows that the error terms are approximately normally distributed, but
the residuals vs. fitted values plot indicates the nonconstancy of the error terms.
As the fitted values increase, the variability of residuals also increases a lot.
So, we will try to find a proper transformation using the Box-Cox procedure below.

```{r P3_2}
par(mfrow=c(1,2))
boxcox(ozone.reg,lambda=seq(-1,2,by=.1))
boxcox(ozone.reg,lambda=seq(0.26,0.3,by=.01))
```

The zoom-in version of the Box-Cox plot on the right shows that $\lambda$ is around 0.28.
Since $\lambda$=0 suggests the logarithmic transformation and $\lambda$=0.5 suggests the
square root transformation of Y, it is not very clear which of the two is more appropriate
in our case. Maybe either one is justifiable. Further analysis seems necessary.

```{r P3_3, fig.height=4}
par(mfrow=c(1,2))
ozone.reg.log <- lm(log(O3) ~ temp+humidity+ibh, data=ozone)
plot(ozone.reg.log$fitted.values, ozone.reg.log$residuals, xlab="Fitted Values", ylab="Residuals",
     main="Log Transformation")
error.std4 <- rstandard(ozone.reg.log)
qqnorm(error.std4, xlab="Normal Scores", ylab="Standardized Residuals", 
       main="Normal Probability Plot")
qqline(error.std4, col="steelblue", lwd=2)

par(mfrow=c(1,2))
ozone.reg.sqrt <- lm(sqrt(O3) ~ temp+humidity+ibh, data=ozone)
plot(ozone.reg.sqrt$fitted.values, ozone.reg.sqrt$residuals, xlab="Fitted Values", ylab="Residuals",
     main="Square Root Transformation")
error.std5 <- rstandard(ozone.reg.sqrt)
qqnorm(error.std5, xlab="Normal Scores", ylab="Standardized Residuals", 
       main="Normal Probability Plot")
qqline(error.std5, col="steelblue", lwd=2)

sh1 <- shapiro.test(ozone.reg$residuals)$p.value
sh2<- shapiro.test(ozone.reg.log$residuals)$p.value
sh3 <- shapiro.test(ozone.reg.sqrt$residuals)$p.value
tbl2 <- data.frame("Original"=sh1, "Log Transformation"=sh2, "Sqrt Transformation"=sh3)
kable(tbl2, format="latex", digits=6, caption="Shapiro-Wilk test p-values") %>%
  kable_styling(latex_options="HOLD_position")

r2 <- summary(ozone.reg)$r.squared
r2.log <- summary(ozone.reg.log)$r.squared
r2.sqrt <- summary(ozone.reg.sqrt)$r.squared
tbl <- data.frame("R-squared"=r2, "R-squared (log)"=r2.log, "R-squared (Sqrt)"=r2.sqrt)
kable(tbl, format="latex", digits=4, caption="Comparison of R-squared Values") %>%
  kable_styling(latex_options="HOLD_position")
```

The QQ plot of the log transformed model has a curvy shape, and the Shapiro-Wilk test
result shows that it is not normally distributed whereas the normality holds for the
square root transformation. 
The residuals plots show that the square root transformation is better than the log
transformation, showing more constancy of variability. Also, $R^2$ is the highest
for the square root transformation despite the small difference. Given above results, 
we would choose the square root transformation, but some additional adjustment might
be necessary because there seems to be significantly lower variances of residuals 
for lower fitted values.
\newpage

## Problem 4

a-)
```{r P4_a}
CP.data <- read.csv("/cloud/project/Data/Commercial Properties Data Set.csv")
plot(CP.data)
round(cor(CP.data),2)
```
The scatter plot matrix and correlation matrix above indicate that each of the predictor
variables is linearly associated with the response variable (Y) with $X_4$ showing the 
highest degrees of association and $X_3$ the lowest. The matrices also show the 
intercorrelations among the predictor variables. In particular, $X_2$ has some degree of
pairwise correlations with other predictor variables while $X_3$ and $X_4$ are 
almost uncorrelated.

b-)
```{r P4_b}
CP.reg <- lm(Y ~ X1+X2+X3+X4, data=CP.data)
summary(CP.reg)
b0 <- CP.reg$coefficients[1] # 12.2
b1 <- CP.reg$coefficients[2] # -0.142
b2 <- CP.reg$coefficients[3] # 0.282
b3 <- CP.reg$coefficients[4] # 0.619
b4 <- CP.reg$coefficients[5] # 0.0000079
```
The estimated regression function:
Y = `r round(b0,3)` - `r round(-b1,3)`$X_1$ + `r round(b2,3)`$X_2$ + `r round(b3,3)`$X_3$ + `r round(b4,7)`$X_4$

c-)
```{r P4_c}
ei <- CP.reg$residuals
boxplot(ei)
```

The box plot of the residuals shows a fairly symmetrical distribution.

d-)
```{r P4_d}
par(mfrow=c(3,2), mar=c(4,4,1,2))
plot(CP.reg$fitted.values, ei, xlab=expression(hat(Y)), ylab="Residuals")
plot(CP.data$X1, ei, xlab=expression(X[1]), ylab="Residuals")
plot(CP.data$X2, ei, xlab=expression(X[2]), ylab="Residuals")
plot(CP.data$X3, ei, xlab=expression(X[3]), ylab="Residuals")
plot(CP.data$X4, ei, xlab=expression(X[4]), ylab="Residuals")
par(mfrow=c(3,2), mar=c(4,4,1,2))
plot(CP.data$X1*CP.data$X2, ei, xlab="X1X2", ylab="Residuals")
plot(CP.data$X1*CP.data$X3, ei, xlab="X1X3", ylab="Residuals")
plot(CP.data$X1*CP.data$X4, ei, xlab="X1X4", ylab="Residuals")
plot(CP.data$X2*CP.data$X3, ei, xlab="X2X3", ylab="Residuals")
plot(CP.data$X2*CP.data$X4, ei, xlab="X2X4", ylab="Residuals")
plot(CP.data$X3*CP.data$X4, ei, xlab="X3X4", ylab="Residuals")
```
```{r fig.height=4}
par(mfrow=c(1,1))
error.std <- rstandard(CP.reg)
qqnorm(error.std, xlab="Normal Scores", ylab="Standardized Residuals", 
       main="Normal Probability Plot")
qqline(error.std, col="steelblue", lwd=2)
```

All the residuals plots do not seem to show any significant pattern. Some of them 
have points concentrated on the left side near zero of the x-axis, but it is 
caused by the differences of scale in the predictor variables. For example, 
30 data points of $X_3$ are zero, so the two-factor interaction terms including
$X_3$ will also have 30 zeros. The normal probability plot indicates that 
the error terms are approximately normally distributed, slightly heavy-tailed.

e-)
```{r P4_e}
bf.data <- data.frame(cbind(CP.data, ei, CP.reg$fitted.values))
dimnames(bf.data)[[2]][6:7] <- c("residuals","fitted.values")
bf.data1<-data.frame(cbind(
  bf.data,ind=as.factor(I(bf.data$fitted.values<median(bf.data$fitted.values))*1)))
g1 <- sum(bf.data1$ind=="1") # group 1: 40 cases
g2 <- sum(bf.data1$ind=="0") # group 2: 41 cases
bf.test(residuals~ind, data=bf.data1)
```
Following the Brown-Forsythe test result shown above, we conclude $H_0$ that
the error variances are constant.