---
title: "Assignment 3"
author: "Chul Hee Kim"
date: '2021 9 26 '
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r include=FALSE, warning=FALSE, message=FALSE}
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(MASS)) install.packages("MASS", repos = "http://cran.us.r-project.org")
if(!require(ALSM)) install.packages("ALSM", repos = "http://cran.us.r-project.org")
if(!require(knitr)) install.packages("knitr", repos = "http://cran.us.r-project.org")
if(!require(kableExtra)) install.packages("kableExtra", repos = "http://cran.us.r-project.org")
if(!require(onewaytests)) install.packages("onewaytests", repos = "http://cran.us.r-project.org")
library(tidyverse)
library(caret)
library(MASS)
library(ALSM)
library(knitr)
library(kableExtra)
library(onewaytests)
```

## Problem 1

a-)
```{r P1_a}
PH.data <- read.csv("/cloud/project/Data/Plastic Hardness Data.csv")
PH.reg <- lm(Y ~ X, data=PH.data)
ei <- PH.reg$residuals
plot(PH.reg$fitted.values, ei, xlab="Fitted Values", ylab="Residuals", main="Plastic Hardness Data")
```

The plot above does not indicate unequal variances, so no departures
from regression model are evident here.

b-)
```{r P1_b}
PHerror.std <- rstandard(PH.reg)
PHqq <- qqnorm(PHerror.std, xlab="Normal Scores", ylab="Standardized Residuals")
qqline(PHerror.std, col="steelblue", lwd=2)
cc <- cor(PHqq$x, PHqq$y) #0.991
```
The normality assumption appears to be reasonable as the above normal
probability plot indicates. The coefficient of correlation between
residuals and their expected values under normality is `r round(cc,3)`,
which also supports our claim.

c-)
```{r P1_c}
std <- 3.234 #from Residual standard error in summary(PH.reg)

# Expected frequencies
p25 <- qt(0.25, 14) * std # -2.239
p50 <- qt(0.5, 14) * std # 0
p75 <- qt(0.75, 14) * std # 2.239

# Residuals and expected values table
tbl <- data.frame(residual=ei)
tbl <- tbl %>% 
  arrange(residual) %>% 
  mutate(rank=seq(1,16)) %>%
  mutate(expected=round(std*qnorm((rank-.375)/(16+.25)),3))  

obs25 <- tbl %>% filter(residual <= p25) %>% count() # 4
obs50 <- tbl %>% filter(residual <= p50) %>% count() # 7
obs75 <- tbl %>% filter(residual <= p75) %>% count() # 11

kable(tbl, caption="Residuals and Expected Values under Normality", "latex") %>%
  kable_styling(position="center", latex_options="hold_position")
```

| Percentiles | t-value          | Expected Freq. | Observed Freq.     |
|-------------|------------------|----------------|--------------------|
| 25th        | `r round(p25,3)` | 4              | `r round(obs25,3)` |
| 50th        | `r round(p50,3)` | 8              | `r round(obs50,3)` |
| 75th        | `r round(p75,3)` | 12             | `r round(obs75,3)` |

Since n = 16, the 25th, 50th, and 75th percentiles of the expected
frequencies under normality are 4, 8, and 12. The t-value for the
corresponding percentiles are -2.239, 0, and 2.239, and the table above
shows that the observed frequencies are 4, 7, and 11. The observed and
expected frequencies are very close to each other, and it is consistent
with part (b) that the normality assumption holds. Also, the residuals and 
their expected values under normality are close to each other.

d-)

```{r P1_d}
# Brown-Forsythe Test
sep <- data.frame(cbind(PH.data, ei))
p1 <- sep[sep[,2] <= 24,]
p2 <- sep[sep[,2] > 24,]
M1 <- summary(p1[,3])[3]
M2 <- summary(p2[,3])[3]
N1 <- dim(p1)[1]
N2 <- dim(p2)[1]

d1 <- abs(p1[,3]-M1)
d2 <- abs(p2[,3]-M2)
s2 <- sqrt((var(d1)*(N1-1)+var(d2)*(N2-1))/(N1+N2-2))

Den <- s2*sqrt(1/N1+1/N2)
Num <- mean(d1)-mean(d2)

T <- Num/Den # 0.856
T.crit <- qt(0.995, N1+N2-2) # 2.977
pvalue <- 2*(1-pt(T, N1+N2-2)) # 0.407
```
\newpage
$H_0$: The error variances are constant. $H_a$: The error variances are
not constant.

Since $t_{BF}^*$ = `r round(T,3)` $\le$ `r round(T.crit,3)`
$(\alpha=0.01)$, we conclude $H_0$, so the error variance is constant
and does not vary with the level of X. The p-value of
`r round(pvalue,3)` $\ge \alpha=0.01$ also supports our claim. This conclusion is
consistent with our preliminary findings in part (a).

```{r}
# Using onewaytests library
bf.data<-data.frame(cbind(PH.data,PH.reg$residuals,PH.reg$fitted.values))
dimnames(bf.data)[[2]][3:4]<-c("residuals","fitted.values")
bf.data1<-data.frame(cbind(bf.data,ind=as.factor(I(bf.data$X<=median(bf.data$X))*1)))
bf.test(residuals~ind,data=bf.data1)
```
The Brown-Forsythe test using 'onewaytests' library gives the same result that
we conclude $H_0$ although the p-value is different from the one above calculated manually
as 'onewaytests' uses F-test, rather than t-test.

e-)
```{r P1_e}
# Breusch-Pagan Test
ei2 <- ei^2
g <- lm(ei2 ~ PH.data$X)
anova(g)
anova(PH.reg)
SSR.S <- anova(g)$`Sum Sq`[1] #66.385
SSE <- anova(PH.reg)$`Sum Sq`[2] #146.425
ChiSquare <- (SSR.S/2) / ((SSE/nrow(PH.data))^2) #0.396
Chi.crit <- qchisq(0.99,1) #6.635
pvalue2 <- 1- pchisq(ChiSquare,1) #0.529
```
$H_0: \gamma_1 = 0$ $H_a: \gamma_1 \ne 0$

Since $\chi_{BP}^2$ = `r round(ChiSquare,3)` $\le$ `r round(T.crit,3)` = 
$\chi^2(0.99,1)$, we conclude $H_0$, so the error variance is constant.
The p-value of `r round(pvalue2,3)` $\ge \alpha=0.01$ also supports our claim. This
conclusion is consistent with our preliminary findings in part (a).

## Problem 2

a-)

```{r P2_a}
SG.data <- read.csv("/cloud/project/Data/Sales Growth Data.csv")

band1 <- subset(SG.data, -0.5<X & X<1.5)
band2 <- subset(SG.data, 1.5<X & X<3.5)
band3 <- subset(SG.data, 3.5<X & X<5.5)
band4 <- subset(SG.data, 5.5<X & X<7.5)
band5 <- subset(SG.data, 7.5<X & X<9.5)

median.X <- c(median(band1$X),median(band2$X),median(band3$X),median(band4$X),median(band5$X))
median.Y <- c(median(band1$Y),median(band2$Y),median(band3$Y),median(band4$Y),median(band5$Y))

plot(SG.data$X, SG.data$Y, xlab="X", ylab="Y", main="Sales Growth Data")
lines(median.X, median.Y)
```

Although there is a slight kink at around X=6.5, the band smooth
suggests that the regression relation is mostly linear.

b-)

```{r P2_b}
band1.1 <- subset(SG.data, -0.5<X & X<2.5)
band2.1 <- subset(SG.data, 0.5<X & X<3.5)
band3.1 <- subset(SG.data, 1.5<X & X<4.5)
band4.1 <- subset(SG.data, 2.5<X & X<5.5)
band5.1 <- subset(SG.data, 3.5<X & X<6.5)
band6.1 <- subset(SG.data, 4.5<X & X<7.5)
band7.1 <- subset(SG.data, 5.5<X & X<8.5)
band8.1 <- subset(SG.data, 6.5<X & X<9.5)

fit1 <- lm(Y ~ X, data=band1.1)
fit2 <- lm(Y ~ X, data=band2.1)
fit3 <- lm(Y ~ X, data=band3.1)
fit4 <- lm(Y ~ X, data=band4.1)
fit5 <- lm(Y ~ X, data=band5.1)
fit6 <- lm(Y ~ X, data=band6.1)
fit7 <- lm(Y ~ X, data=band7.1)
fit8 <- lm(Y ~ X, data=band8.1)

M1X <- median(band1.1$X)
M2X <- median(band2.1$X)
M3X <- median(band3.1$X)
M4X <- median(band4.1$X)
M5X <- median(band5.1$X)
M6X <- median(band6.1$X)
M7X <- median(band7.1$X)
M8X <- median(band8.1$X)

Y1 <- predict(fit1,data.frame(X=M1X))
Y2 <- predict(fit2,data.frame(X=M2X))
Y3 <- predict(fit3,data.frame(X=M3X))
Y4 <- predict(fit4,data.frame(X=M4X))
Y5 <- predict(fit5,data.frame(X=M5X))
Y6 <- predict(fit6,data.frame(X=M6X))
Y7 <- predict(fit7,data.frame(X=M7X))
Y8 <- predict(fit8,data.frame(X=M8X))

Xc <- c(M1X,M2X,M3X,M4X,M5X,M6X,M7X,M8X)
Yc_hat <- c(Y1,Y2,Y3,Y4,Y5,Y6,Y7,Y8)

plot(SG.data$X, SG.data$Y, xlab="X", ylab="Y", main="Sales Growth Data")
lines(Xc, Yc_hat)
```

c-)

```{r P2_c}
true.reg <- lm(Y ~ X, data=SG.data)
bd<-predict(true.reg,SG.data,se.fit=TRUE)

W=sqrt(2*qf(0.95,2,bd$df))

CB_lower<-bd$fit-W*bd$se.fit
CB_upper<-bd$fit+W*bd$se.fit
dat<-data.frame(X=SG.data$X,CB_lower,CB_upper)

plot(SG.data$X, SG.data$Y, xlab="X", ylab="Y", main="Sales Growth Data")
lines(Xc, Yc_hat)
lines(dat$X,dat$CB_lower,lty=1,lwd=2, col="steelblue")
lines(dat$X,dat$CB_upper,lty=1,lwd=2, col="steelblue")
```

The simplified lowess smooth falls entirely within the confidence band
for the regression line, and it shows that the linear regression is
appropriate.

## Problem 3

a-)

```{r P3_a}
PT.data <- read.csv("/cloud/project/Data/Production Time.csv")
plot(PT.data$X, PT.data$Y, xlab="X", ylab="Y", main="Production Time Data")
```

A linear relation does not seem adequate here since the scatter plot
above shows a curvilinear shape. The transformation on X would be more
appropriate in case of nonlinear relation, which is the case here. In
particular, 'log' or 'square root' transformation seems appropriate,
given the pattern of the scatter plot. 

b-)

```{r P3_b}
trans.X <- sqrt(PT.data$X)
trans.PT.data <- data.frame(cbind(PT.data, trans.X))
trans.PT.reg <- lm(Y ~ trans.X, data=trans.PT.data)
b0.p3 <- trans.PT.reg$coefficients[1]
b1.p3 <- trans.PT.reg$coefficients[2]
summary(trans.PT.reg)
```

The estimated linear regression function for the transformed data: \newline
$X'$ = $\sqrt{X}$ \newline
Y = `r round(b0.p3,3)` + `r round(b1.p3,3)`$X'$

c-)

```{r P3_c}
plot(trans.PT.data$trans.X, trans.PT.data$Y, xlab=expression(sqrt(X)), ylab="Y",
     main="(Transformed) Production Time Data")
abline(trans.PT.reg)
```

The regression line appear to be a good fit to the transformed data as
the plot above shows.

d-)

```{r P3_d, fig.height=3.5}
par(mfrow=c(1,2))
ei.p3 <- trans.PT.reg$residuals
fitted.p3 <- trans.PT.reg$fitted.values
plot(fitted.p3, ei.p3, xlab="Fitted Values", ylab="Residuals", main="(Transformed) Production Time Data", cex.main=0.7)

PTerror.std <- rstandard(trans.PT.reg)
qqnorm(PTerror.std, xlab="Normal Scores", ylab="Standardized Residuals")
qqline(PTerror.std, col="steelblue", lwd=2)
```

The residuals vs. fitted values plot on the left shows no noticeable
pattern, so the error terms seem to have constant variances. The normal
probability plot on the right shows that it is approximately normally
distributed.

e-) Y = `r round(b0.p3,3)` + `r round(b1.p3,3)`$\sqrt{X}$

## Problem 4

a-)

```{r P4_a}
# SG.data and true.reg already exist from P2
plot(SG.data$X, SG.data$Y, xlab="X", ylab="Y", main="Sales Growth Data")
par(mfrow=c(1,2))
boxcox(true.reg, lambda=seq(-1,2,by=.1))

# Evaluate SSE
boxcox.sse(SG.data$X, SG.data$Y, l=seq(.3,.7,by=.1))
```

The scatter plot shows that a linear relation appears adequate, but the
Box-Cox plot indicates $\lambda = 0.5$, which means the square root
transformation of Y is suggested. The table above also shows that SSE is
minimized when $\lambda = 0.5$. We can infer from this fact needing the
transformation of Y that either the error terms do not have constant variances,
or it is not normally distributed.

b-)

```{r P4_b}
trans.Y <- sqrt(SG.data$Y)
trans.SG.data <- data.frame(SG.data, trans.Y)
trans.SG.reg <- lm(trans.Y ~ X, data=trans.SG.data)
b0.p4 <- trans.SG.reg$coefficients[1] #10.261
b1.p4 <- trans.SG.reg$coefficients[2] #1.076
summary(trans.SG.reg)
```

The estimated linear regression function for the transformed data: \newline
$Y'$ = $\sqrt{Y}$ \newline
$Y'$ = `r round(b0.p4,3)` + `r round(b1.p4,3)`X

c-)

```{r P4_c, fig.height=10}
par(mfrow=c(3,2))
# Regression lines
plot(SG.data$X, SG.data$Y, xlab="X", ylab="Y", main="Sales Growth Data")
abline(true.reg)

plot(trans.SG.data$X, trans.SG.data$trans.Y, xlab="X", ylab=expression(sqrt(Y)),
     main="(Transformed) Sales Growth Data")
abline(trans.SG.reg)

# Residuals vs. fitted values
ei.p41 <- true.reg$residuals
fitted.p41 <- true.reg$fitted.values
plot(fitted.p41, ei.p41, xlab="Fitted Values", ylab="Residuals", main="Sales Growth Data")

ei.p42 <- trans.SG.reg$residuals
fitted.p42 <- trans.SG.reg$fitted.values
plot(fitted.p42, ei.p42, xlab="Fitted Values", ylab="Residuals", 
     main="(Transformed) Sales Growth Data")

# QQ plot
SGerror.std1 <- rstandard(true.reg)
qqnorm(SGerror.std1, xlab="Normal Scores", ylab="Standardized Residuals", 
       main="QQ Plot - Sales Growth Data")
qqline(SGerror.std1, col="steelblue", lwd=2)

SGerror.std2 <- rstandard(trans.SG.reg)
qqnorm(SGerror.std2, xlab="Normal Scores", ylab="Standardized Residuals", 
       main="QQ Plot - (Transformed) Sales Growth Data")
qqline(SGerror.std2, col="steelblue", lwd=2)
```

The scatter plots on the top indicate that both the original and
transformed data appear to be a good fit to the corresponding linear
regression although the transformed model has a better fit.

The residuals vs. fitted values plots in the middle indicate that the
variances of error terms are not constant in the original data whereas
those seem constant in the transformed data.

The normal probability plots on the bottom indicate that both the
original and transformed data are approximately normally distributed,
but we can notice a slightly curvy shape in the plot of the original
data.

d-) $\sqrt{Y}$ = `r round(b0.p4,3)` + `r round(b1.p4,3)`X

## Problem 5

a-)

```{r P5_a}
X <- c(5,8,11,7,13,12,12,6) #diastolic blood pressure
Y <- c(63,67,74,64,75,69,90,60) #age for boys 5 to 13
p5.data <- data.frame(cbind(X, Y))
p5.reg <- lm(Y ~ X, data=p5.data)
b0.p5 <- p5.reg$coefficients[1] #48.667
b1.p5 <- p5.reg$coefficients[2] #2.333

ei.p5 <- p5.reg$residuals
plot(p5.data$X, ei.p5, xlab="X", ylab="Residuals", main="Diastolic Blood Pressure (Y) vs. Age for boys 5-13 (X)")
```

The residual plot above shows that the error terms do not have constant
variances and seem to have larger variances where X is bigger.

The estimated regression function: Y = `r round(b0.p5,3)` +
`r round(b1.p5,3)`X
\newpage
b-)

```{r P5_b}
p5.newdata <- p5.data[-7,]
p5.newreg <- lm(Y ~ X, data=p5.newdata)
b0.p5new <- p5.newreg$coefficients[1] #53.068
b1.p5new <- p5.newreg$coefficients[2] #1.621
```

The estimated regression function: Y = `r round(b0.p5new,3)` +
`r round(b1.p5new,3)`X

After omitting the case 7 from the data, the regression function has
smaller $b_1$, so the case 7 affected the slope of the initial regression function
to be steeper. We could have expected it by seeing the original data because Y=90
at case 7 whereas Y sits 60 ~ 75 for all other cases. The case 7 seems to be an outlier.

c-)

```{r P5_c}
predict.int <- predict(p5.newreg, data.frame(X=12), interval="prediction", level=0.99)
lwr <- predict.int[2] #60.313
upr <- predict.int[3] #84.736
```

99% Prediction interval for a new Y observation at X=12:
`r round(lwr,3)` $\le Y_{h(new)} \le$ `r round(upr,3)`

The observation $Y_6$ falls inside this prediction interval. Falling
inside the prediction interval is not very meaningful as it is designed
to be rather wide and to contain most of the observations at X=12. 

If the observation actually fell outside the prediction interval as the observation
$Y_7$ does, we should be alerted that maybe the blood pressure was measured incorrectly,
or that 12-year boy has something going on in his body, etc.
