---
title: "Assignment 7"
author: "Chul Hee Kim"
date: '2021 10 31 '
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("scipen" = 10)  ##Used to avoid displaying numbers such as 1+e3 (exponent), changing the options
# STEP 0: Function to load or download libraries
loadLib = function(libName)
{
  if(require(libName, character.only=TRUE))
  {
    ##cat(libName, "loaded properly\n")
  } else {
    cat("Installing", libName, "\n")
    install.packages(libName)
    if(require(libName, character.only=TRUE))
    {
      cat(libName, "loaded properly\n")
    } else {
      stop(c(libName, "not properly installed\n"))
    }
  }
}

libs = c("ggplot2","knitr","MASS","ggpubr","dplyr","tidyverse","faraway", 
         "rmarkdown", "markdown", "tidyr", "caret","readr","onewaytests","plyr",
         "lmtest","pander","kableExtra","QuantPsyc","GGally","olsrr","leaps",
         "ggcorrplot","glmnet")

for (lib in libs)
{
  loadLib(lib)
}

## Helper to get data sets from one location on my R Cloud project
getDataSet <- function(dataset) {
  datasets.loc <- "/cloud/project/Data/"
  read.csv(paste0(datasets.loc,dataset,sep=""))
}
```

## Problem 1

a-)
```{r P1_a}
CCdata <- getDataSet("Cement Composition.csv")
f1 <- lm(Y ~ ., data=CCdata)
summary(f1)
vif(f1)
```
Y = 62.405 + 1.551$X_1$ + 0.510$X_2$ + 0.102$X_3$ - 0.144$X_4$

$R^2$ is very big, and the model is significant but none of the predictor variables 
are significant with $\alpha$=0.05. VIFs indicate multicollinearity problem.

b-)
```{r P1_b}
x <- model.matrix(Y~., CCdata)[,-c(1)]
y <- CCdata$Y
RidgeMod <- glmnet(x, y, alpha=0, nlambda=100,lambda.min.ratio=0.0001)
CvRidgeMod <- cv.glmnet(x, y, alpha=0, nlambda=100,lambda.min.ratio=0.0001)

plot(CvRidgeMod)
best.lambda.ridge <- CvRidgeMod$lambda.min
pred.ridge <- coef(CvRidgeMod, s = "lambda.min")
pred.ridge
```
The best $\lambda$ for the ridge regression model is `r round(best.lambda.ridge,4)`

c-)
```{r P1_c}
LassoMod <- glmnet(x, y, alpha=1, nlambda=100,lambda.min.ratio=0.0001)
CvLassoMod <- cv.glmnet(x, y, alpha=1, nlambda=100,lambda.min.ratio=0.0001)
plot(CvLassoMod)
best.lambda.lasso <- CvLassoMod$lambda.min
pred.lasso <- coef(CvLassoMod, s = "lambda.min")
pred.lasso
```
The best $\lambda$ for the lasso regression model is `r round(best.lambda.lasso,4)`.
This model gets rid of $X_3$.

d-)
```{r P1_d}
EnetMod <- glmnet(x, y, alpha=0.5, nlambda=100,lambda.min.ratio=0.0001)
CvElasticnetMod <- cv.glmnet(x, y,alpha=0.5,nlambda=100,lambda.min.ratio=0.0001)
best.lambda.enet <- CvElasticnetMod$lambda.min
pred.enet <- coef(CvElasticnetMod, s = "lambda.min")
pred.enet
```
The best $\lambda$ for the elastic net model is `r round(best.lambda.enet,4)`

e-)
```{r P1_e}
yhat.ridge <- predict(RidgeMod, s = best.lambda.ridge, newx = x)
yhat.lasso <- predict(LassoMod, s = best.lambda.lasso, newx = x)
yhat.enet <- predict(EnetMod , s = best.lambda.enet, newx = x)
sst <- sum((y - mean(y))^2)
sse.ols<-sum(f1$residuals^2)
sse.ridge <- sum((y - yhat.ridge)^2)
sse.lasso <- sum((y - yhat.lasso)^2)
sse.enet <- sum((y - yhat.enet)^2)
tbl.sse <- cbind(sse.ols,sse.ridge,sse.lasso,sse.enet)
kable(tbl.sse, digits=4, align="c", caption="SSE Comparison", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
# R squared
rsq.ols <- 1 - sse.ols / sst
rsq.ridge <- 1 - sse.ridge / sst
rsq.lasso <- 1 - sse.lasso / sst
rsq.enet  <- 1 - sse.enet  / sst
tbl.r2 <- cbind(rsq.ols,rsq.ridge,rsq.lasso,rsq.enet)
kable(tbl.r2, digits=4, align="c", caption="R Squared Comparison", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
SSE is the smallest for the ordinary least squares model (OLS), but the VIFs shown
in part (a) indicate that there is multicollinearity issue. The optimal model of 
our selection would be the lasso regression model because it has the next smallest
SSE, but the elastic net model comes pretty close. The lasso model is also preferred
over the elastic net model since the lasso model uses one fewer predictor variable.


## Problem 2

a-)
```{r P2_a}
data(stackloss)
f2 <- lm(stack.loss ~ ., data=stackloss)
summary(f2)
vif(f2)
```
$R^2$ is very big, and the model is significant. One predictor variable,
Acid.Conc., looks not significant, and the VIFs show no multicollinearity problem.

b-)
```{r P2_b}
rr.huber <- rlm(stack.loss ~ ., data = stackloss, psi = psi.huber)
summary(rr.huber)
```

c-)
```{r P2_c}
rr.bisquare <- rlm(stack.loss ~ ., data=stackloss, psi = psi.bisquare)
summary(rr.bisquare)
```

d-)
```{r P2_d1}
tbl.robust <- cbind(f2$coefficients, rr.huber$coefficients, rr.bisquare$coefficients)
colnames(tbl.robust) <- c("OLS", "Huber", "Bisquare")
kable(tbl.robust, digits=4, align="c", caption="OLS vs. Robust Comparison", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
Because the coefficients for OLS are quite different from those for robust regression
models, especially for Water.Temp, we would use the robust regression models. It is 
an indication that there exist outliers or high leverage points.

```{r P2_d2, fig.height=4}
par(mfrow=c(2,2))
ols_plot_cooksd_chart(f2)
ols_plot_resid_stud(f2)
ols_plot_resid_lev(f2)
ols_plot_resid_stud_fit(f2)
```
The four diagnostic plots show that the data point '21' could be an outlier.
'4' is also identified as an outlier, but we will keep it in the model since
it is on the border line. Also, the data set is quite small with only 21 points.
We will also remove '17' for its high leverage.


```{r P2_d3}
stackloss2 <- stackloss[-c(17,21),]
f2.1 <- lm(stack.loss ~ ., data=stackloss2)
summary(f2.1)

tbl.robust2 <- cbind(f2.1$coefficients, rr.huber$coefficients, rr.bisquare$coefficients)
colnames(tbl.robust2) <- c("OLS(New)", "Huber", "Bisquare")
kable(tbl.robust2, digits=4, align="c", caption="OLS vs. Robust Comparison", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
After removing an outlier and a high leverage point, we find that the new OLS has 
a lot similar coefficients to those of the robust regression models. 


## Problem 3

a-) Linear Regression
```{r P3_a}
data(fat)
fat.data <- fat[,-c(1,3)]

set.seed(1023)
train.index <- createDataPartition(y = fat.data$siri, p = 0.7, list = FALSE)
train.set <- fat.data[train.index,]
test.set <- fat.data[-train.index,]

f3 <- lm(siri ~ ., data=train.set)
stepwise.result <- ols_step_both_p(f3, pent=0.1, prem=0.05)
stepwise.result

f3.ols <- lm(siri ~ abdom+free+weight+adipos+biceps+thigh+age+neck, data=train.set)
summary(f3.ols)
```

c-) Ridge Regression
```{r P3_c}
x3 <- model.matrix(siri~., train.set)[,-c(1)]
y3 <- train.set$siri
RidgeMod3 <- glmnet(x3, y3, alpha=0, nlambda=100,lambda.min.ratio=0.0001)
CvRidgeMod3 <- cv.glmnet(x3, y3, alpha=0, nlambda=100,lambda.min.ratio=0.0001)
par(mfrow=c(1,1))
plot(CvRidgeMod3)
best.lambda.ridge3 <- CvRidgeMod3$lambda.min
pred.ridge3 <- coef(CvRidgeMod3, s = "lambda.min")
pred.ridge3
```

d-) Lasso
```{r P3_d}
LassoMod3 <- glmnet(x3, y3, alpha=1, nlambda=100,lambda.min.ratio=0.0001)
CvLassoMod3 <- cv.glmnet(x3, y3, alpha=1, nlambda=100,lambda.min.ratio=0.0001)
plot(CvLassoMod3)
best.lambda.lasso3 <- CvLassoMod3$lambda.min
pred.lasso3 <- coef(CvLassoMod3, s = "lambda.min")
pred.lasso3
```

e-) Elastic Net
```{r P3_e}
EnetMod3 <- glmnet(x3, y3, alpha=0.5, nlambda=100,lambda.min.ratio=0.0001)
CvElasticnetMod3 <- cv.glmnet(x3, y3,alpha=0.5,nlambda=100,lambda.min.ratio=0.0001)
plot(CvElasticnetMod3)
best.lambda.enet3 <- CvElasticnetMod3$lambda.min
pred.enet3 <- coef(CvElasticnetMod3, s = "lambda.min")
pred.enet3
```

f-) Robust Regression
```{r P3_f}
rr.huber3 <- rlm(siri ~ ., data = train.set, psi = psi.huber)
summary(rr.huber3)

rr.bisquare3 <- rlm(siri ~ ., data=train.set, psi = psi.bisquare)
summary(rr.bisquare3)
```

g-)
```{r P3_g}
xt <- model.matrix(siri~., test.set)[,-c(1)]
yhat.ridge3 <- predict(RidgeMod3, s = best.lambda.ridge3, newx = xt)
yhat.lasso3 <- predict(LassoMod3, s = best.lambda.lasso3, newx = xt)
yhat.enet3 <- predict(EnetMod3, s = best.lambda.enet3, newx = xt)
yhat.huber <- predict(rr.huber3, test.set)
yhat.bisquare <- predict(rr.bisquare3, test.set)

y4 <- test.set$siri
sse.ols3 <- sum(f3.ols$residuals^2)
sse.ridge3 <- sum((y4 - yhat.ridge3)^2)
sse.lasso3 <- sum((y4 - yhat.lasso3)^2)
sse.enet3 <- sum((y4 - yhat.enet3)^2)
sse.huber <- sum((y4 - yhat.huber)^2)
sse.bisquare <- sum((y4 - yhat.bisquare)^2)

tbl.sse3 <- cbind(sse.ols3,sse.ridge3,sse.lasso3,sse.enet3,sse.huber,sse.bisquare)
colnames(tbl.sse3) <- c("OLS","Ridge","Lasso","ENet","Huber","Bisquare")
kable(tbl.sse3, digits=3, align="c", caption="SSE Comparison", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
The Huber model has the smallest SSE and would be the best candidate for the final
selection although the Lasso and Bisquare models' performances are also not bad.
The OLS also has the similar SSE as the Huber, but VIFs shown below indicate 
multicollinearity issue.

```{r}
vif(f3.ols)
```

## Problem 4

a-)
```{r P4_a}
MSdata <- getDataSet("Machine Speed.csv")
f4 <- lm(Y ~ X, data=MSdata)
summary(f4)
ei <- f4$residuals
par(mfrow=c(1,2))
plot(MSdata$X, ei, xlab="X", ylab="Residuals")
plot(MSdata$X, ei^2, xlab="X", ylab="Squared Residuals")
```

The plots above show a megaphone shape, indicating that the error variances 
increase as X gets larger. But, since the size of the data set is quite small,
we would also want to check the significance with some formal test.

b-)
```{r P4_b}
bptest(f4)
```
$H_0$: The error variances are constant. \
$H_a$: The error variances are not constant. 

The p-value of Breusch-Pagan test is smaller than $\alpha$=0.1, so we conclude
$H_a$ that the error variances are not constant. It is consistent with what we observed
with residual plots in part (a).

c-)
```{r P4_c}
abs.ei <- abs(ei)
g1 <- lm(abs.ei ~ MSdata$X)
s <- g1$fitted.values
w1 <- 1/(s^2)
#use the weights to fit weighted regression model
g2 <- lm(Y ~ X, weights= w1, data=MSdata)
summary(g2)

tbl.4c <- data.frame(cbind(f4$coefficients, g2$coefficients))
colnames(tbl.4c) <- c("Ordinary","Weighted")
kable(tbl.4c, digits=4, align="c", caption="Ordinary vs. Weighted Least Squares", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
For $\beta_1$, it is almost identical for both models.
For $\beta_0$, it is changed a little bit but does not appear to be significantly different.

e-)
```{r P4_e}
tbl.4e <- data.frame(cbind(summary(f4)$coefficients[,2], summary(g2)$coefficients[,2]))
colnames(tbl.4e) <- c("Ordinary","Weighted")
kable(tbl.4e, digits=4, align="c", caption="Standard Deviations of Estimated Coefficients", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
For $\beta_1$, SD is almost identical for both models.
For $\beta_0$, SD is reduced slightly.

f-)
```{r P4_f}
ei2 <- g2$residuals
abs.ei2 <- abs(ei2)
g3 <- lm(abs.ei2 ~ MSdata$X)
s2 <- g3$fitted.values
w2 <- 1/(s2^2)
#use the weights to fit weighted regression model
g4 <- lm(Y ~ X, weights= w2, data=MSdata)
summary(g4)

tbl.4f <- data.frame(cbind(f4$coefficients, g2$coefficients, g4$coefficients))
colnames(tbl.4f) <- c("Ordinary","Weighted1","Weighted2")
kable(tbl.4f, digits=4, align="c", caption="Ordinary vs. Weighted Least Squares", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
There is no substantial change in the estimated regression coefficients from the first
weighted least squares model. In fact, it is almost the same. If there were a big change,
we should iterate the previous steps until the estimates of coefficients converge. But,
here there is no need for iteration since it seems to be already converged.
