---
title: "Assignment 6"
author: "Chul Hee Kim"
date: '2021 10 25 '
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("scipen" = 10)  ##Used to avoid displaying numbers such as 1+e3 (exponent), changing the options
# STEP 0: Function to load or download libraries
loadLib = function(libName)
{
  if(require(libName, character.only=TRUE))
  {
    ##cat(libName, "loaded properly\n")
  } else {
    cat("Installing", libName, "\n")
    install.packages(libName)
    if(require(libName, character.only=TRUE))
    {
      cat(libName, "loaded properly\n")
    } else {
      stop(c(libName, "not properly installed\n"))
    }
  }
}

libs = c("ggplot2","knitr","MASS","ggpubr","dplyr","tidyverse","faraway", 
         "rmarkdown", "markdown", "tidyr", "caret","readr","onewaytests","plyr",
         "lmtest","pander","kableExtra","QuantPsyc","GGally","olsrr","leaps","ggcorrplot")

for (lib in libs)
{
  loadLib(lib)
}

## Helper to get data sets from one location on my R Cloud project
getDataSet <- function(dataset) {
  datasets.loc <- "/cloud/project/Data/"
  read.csv(paste0(datasets.loc,dataset,sep=""))
}
```

## Problem 1

```{r}
data(prostate)
head(prostate)
str(prostate)
f1 <- lm(lpsa ~ ., data=prostate)
summary(f1)
```

a-) Backward elimination

We will eliminate the largest p-value among those over 0.05, one at a time.
```{r P1_a}
f11 <- update(f1, .~.-gleason)
summary(f11)
f12 <- update(f11, .~.-lcp)
summary(f12)
f13 <- update(f12, .~.-pgg45)
summary(f13)
f14 <- update(f13, .~.-age)
summary(f14)
f15 <- update(f14, .~.-lbph)
summary(f15)
```
The model selected by the backward elimination:\
lpsa = -0.268 + 0.552lcavol + 0.509lweight +0.666svi

Below result confirms the same variables removed by the backward elimination:
```{r}
ols_step_backward_p(f1, prem=0.05)
```

b-) AIC
```{r P1_b}
b <- regsubsets(lpsa ~ ., data=prostate)
rs <- summary(b)
rs$which

AIC<-97*log(rs$rss/97)+(2:9)*2
plot(AIC~I(1:8),ylab="AIC",xlab="Number of Predictors")
which.min(AIC)
```
The plot above shows that AIC is minimized when there are 5 predictors, and the table
suggests that we choose 'lcavol', 'lweight', 'age', 'lbph', and 'svi'.

Below stepwise method confirms the above result.
```{r}
step(f1)
```
lpsa = 0.951 + 0.566lcavol + 0.424lweight - 0.015age + 0.112lbph + 0.721svi

c-) Adjusted $R_a^2$
```{r P1_c}
plot(2:9,rs$adjr2,xlab="Number of Parameters",ylab="Adjusted R Square")
which.max(rs$adjr2)
```
The adjusted $R^2$ is maximized when there are 8 parameters, but the marginal increase
is very small once the number of parameters reaches 5 or 6.

d-) Mallows $C_p$
```{r P1_d, fig.height=4}
plot(2:9,rs$cp,xlab="Number of Parameters",ylab="Cp")
abline(0,1)

ols_step_best_subset(f1)
```
We would choose between the six parameter, five predictor, model (Model 5) including
'lcavol', 'lweight', 'age', 'lbph', and 'svi', and the model also including
'pgg45' (Model 6). Model 7 is also below $C_p$=p line, but we prefer the ones
with fewer predictors. 

Model 5 is the closest to $C_p$=p line and has the smallest AIC, but we would still
keep Model 6, which has similarly good numbers, as another candidate and 
take a deeper analysis (i.e. check assumptions) to select the final model.


## Problem 2

a-)
```{r P2_a, fig.height=4.3}
SENIC <- getDataSet("SENIC.csv")
SENIC.r <- SENIC[,-c(1,8:9)] # drop 'MedicalSchoolAffiliation' and 'Region'

train.set <- SENIC.r[57:113,]
test.set <- SENIC.r[1:56,]

par(mfrow=c(2,2))
dotchart(train.set$Age, xlab="Age")
dotchart(train.set$InfectionRisk, xlab="InfectionRisk")
dotchart(train.set$CulturingRatio, xlab="CulturingRatio")
dotchart(train.set$ChestXRayRatio, xlab="ChestXRayRatio")
dotchart(train.set$Beds, xlab="Beds")
dotchart(train.set$AvgDailyCensus, xlab="AvgDailyCensus")
dotchart(train.set$Nurses, xlab="Nurses")
dotchart(train.set$AvailableFacilitiesPercent, xlab="AvailableFacilitiesPercent")
```

We see some outlying data points (e.g. one below 40 for Age, about 800 for Beds/AvgDailyCensus, 
below 10% for AvailableFacilitiesPercent), but there does not appear to be problematic features in 
the dot plots. It is notable that the variables like 'Beds', 'AvgDailyCensus', and 'Nurses'
have data points concentrated on the left-hand side. 'AvailableFacilitiesPercent' almost looks like
a categorical variable, but it is just due to the fact that there are many hospitals with 
the similar number of facilities and services.

b-)
```{r P2_b}
plot(train.set)
corr <- round(cor(train.set),2)
ggcorrplot(corr, type="lower", lab=TRUE)
```

The following pairs of predictor variables have very strong linear pairwise associations:\
Beds and AvgDailyCensus: r = 0.99\
Beds and Nurses: r = 0.91\
Nurses and AvgDailyCensus: r = 0.90

The following pairs also have quite strong linear pairwise associations:\
AvailableFacilitiesPercent and Beds: r = 0.76\
AvailableFacilitiesPercent and AvgDailyCensus: r = 0.73\
AvailableFacilitiesPercent and Nurses: r = 0.71

c-)
```{r P2_c}
f2 <- lm(log(Stay) ~ ., data=train.set)
ols_step_best_subset(f2)
Cp <- ols_step_best_subset(f2)[3:5,7]
p <- c(4,5,6)
cpp <- data.frame(cbind(Cp, p))
rownames(cpp) <- c("Model 3","Model 4","Model 5")
kable(cpp, digits=3, align="c", caption="Cp and p Comparison", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
According to the table above, Model 3, 4, and 5 are the three best subsets for the $C_p$ criterion.
Among these three subsets, Model 3 appears to have the smallest bias since Model 3
has $C_p$=3.8112 and p=4. The gap is only 0.1888, and Model 4 and 5 have bigger gaps
between $C_p$ and p. Also, we want $C_p$ to be small and to be close to p.

d-)
```{r P2_d}
f2.train <- lm(log(Stay) ~ Age+ChestXRayRatio+AvgDailyCensus, data=train.set)
s1 <- summary(f2.train)
s1

f2.test <- lm(log(Stay) ~ Age+ChestXRayRatio+AvgDailyCensus, data=test.set)
s2 <- summary(f2.test)
s2

b1.co <- c(s1$coefficients[2,1], s2$coefficients[2,1])
b2.co <- c(s1$coefficients[3,1], s2$coefficients[3,1])
b3.co <- c(s1$coefficients[4,1], s2$coefficients[4,1])

b1.sd <- c(s1$coefficients[2,2], s2$coefficients[2,2])
b2.sd <- c(s1$coefficients[3,2], s2$coefficients[3,2])
b3.sd <- c(s1$coefficients[4,2], s2$coefficients[4,2])

tbl1 <- data.frame(rbind(b1.co, b2.co, b3.co))
rownames(tbl1) <- c("b1","b2","b3")
colnames(tbl1) <- c("Train","Test")

tbl2 <- data.frame(rbind(b1.sd, b2.sd, b3.sd))
rownames(tbl2) <- c("b1","b2","b3")
colnames(tbl2) <- c("Train","Test")

kable(tbl1, digits=5, align="c", caption="Regression Coefficients", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
kable(tbl2, digits=5, align="c", caption="Standard Deviations of Regression Coefficients", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
The estimated regression coefficients and their standard deviations do not differ
much between the two models. But, for the validation set, the summary shows that 
the p-value for 'Age' is bigger than 0.05, making it insignificant for 
the regression model using the validation set whereas it is significant for that
using the training set.

e-)
```{r P2_e}
MSE.train <- anova(f2.train)$`Mean Sq`[4]
MSE.test <- anova(f2.test)$`Mean Sq`[4]
tbl3 <- data.frame(MSE.train, MSE.test)

tbl4 <- data.frame(s1$r.squared, s2$r.squared)
colnames(tbl4) <- c("Rsquared(Train)","Rsquared(Test)")

kable(tbl3, digits=5, align="c", caption="MSE", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
kable(tbl4, digits=5, align="c", caption="Coefficients of Multiple Determination", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
MSEs do not differ much, but $R^2$ is significantly reduced for the validation set compared to
the training set. Both models are significant, but considering very different $R^2$ and 
bigger p-value of 'Age' for the model fitted to the validation set, we need a further analysis.
It could be the case that some other subset model fits better to the validation set.


## Problem 3

```{r P3}
data(teengamb)
f3 <- lm(gamble ~ ., data=teengamb)
summary(f3)

stepwise.result <- ols_step_both_p(f3, pent=0.1, prem=0.05)
stepwise.result
summary(stepwise.result$model)
```
gamble = 4.041 + 5.172income - 21.634sex

We see below that adding 'verbal' or 'status' does not make either variable 
significant, so the above final model with only 'income' and 'sex' is confirmed
as the final one, and both variables are significant.
```{r}
f3.2 <- lm(gamble ~ income+sex+verbal, data=teengamb)
summary(f3.2)
f3.3 <- lm(gamble ~ income+sex+status, data=teengamb)
summary(f3.3)
```