---
title: "Assignment 2"
author: "Chul Hee Kim"
date: '2021 9 17 '
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

## Problem 1

```{r P1_setup, message=FALSE}
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
library(tidyverse)
library(caret)
crime_data <- read.csv("/cloud/project/Data/Crime Rate.csv")
```

a-)

```{r P1_a}
fit_crime <- lm(Y ~ X, data = crime_data)
anova(fit_crime)
```

b-)

```{r P1_b1}
n <- nrow(crime_data)
f <- qf(0.95, 1, n-2)
```

$H_0: \beta_1 = 0$ \newline $H_a: \beta_1 \ne 0$ \newline

ANOVA table above shows $F^*$ = 16.834, and $F(1-\alpha;1,n-2)$ =
`r round(f,3)`. Since F-statistic is bigger, we conclude $H_a$, and the
model is significant.

```{r P1_b2}
summary(fit_crime)
t <- qt(0.975,n-2)
```

The summary above shows that the $t^*$ for X is -4.103. The equivalence
of F test and t test says that $F^*$ = $(t^*)^2$, and it is consistent
with above results since

$(t^*)^2 = (-4.103)^2 = 16.83 = F^*$.

$F(1-\alpha;1,n-2)$ = `r round(f,3)` \newline $[t(1-\alpha/2;n-2)]^2$ =
`r round(t^2,3)`

Therefore, we have shown the numerical equivalence of the two test
statistics and decision rules, which should be the case with
one-predictor-variable model. The p-values for the two tests are also
shown in the summary above. The one for t test appears in the row same
as X and reads 9.57e-05. The one for F test appears at the bottom and
reads 9.571e-05, and those two p-values are the same.

c-) As the R-squared value in the summary shown in (b) explains, the
total variation in crime rate is reduced by 17.03% when the percentage
of high school graduates is introduced into the analysis. This is a
relatively small reduction because there is a lot of room for more
reduction by introducing other variables. However, one can still argue
that the percentage of high school graduates is significant, given its
small p-value. The point we want to emphasize here is that the model still 
falls short of predicting the crime rate.

d-)

```{r P1_d}
b0_full <- fit_crime$coefficients[1]
b1_full <- fit_crime$coefficients[2]

fit_crime_red <- lm(Y ~ 1, data = crime_data)
b0_red <- fit_crime_red$coefficients[1]
```

Full Model: $Y_i = \beta_0 + \beta_1X_i + \epsilon_i =$
`r round(b0_full,3)` + `r round(b1_full,3)`X + $\epsilon_i$

Reduced Model: $Y_i = \beta_0 + \epsilon_i =$ `r round(b0_red,3)` +
$\epsilon_i$

e-)

```{r P1_e}
anova(fit_crime)
anova(fit_crime_red)
SSE_F <- anova(fit_crime)$`Sum Sq`[2]
SSE_R <- anova(fit_crime_red)$`Sum Sq`[1]
f_stat <- ((SSE_R-SSE_F)/(83-82)) / (SSE_F/82)
f_gl <- qf(0.95,1,82)
```

(1) SSE(F) = `r SSE_F`
(2) SSE(R) = `r SSE_R`
(3) $df_F$ = n - 2 = 82
(4) $df_R$ = n - 1 = 83
(5) $F^* = \frac{SSE(R)-SSE(F)}{df_R-df_F}\div\frac{SSE(F)}{df_F} = \frac{455,273,165-548,736,108}{83-82}\div\frac{548,736,108}{82}$
    = 16.834
(6) $F(1-\alpha;df_R-df_F,df_f) = F(0.95;1,82)$ = `r round(f_gl,3)`

f-) Yes, the $F^*$ and the decision rule for the general linear test
numerically equivalent to those in part a,b as the results above show.

## Problem 2

```{r P2_setup}
X <- c(4,8,12,16,20)
```

a-)

```{r P2_a}
errors <- rnorm(n=5, mean=0, sd=5)
dat <- data.frame(Y = 20 + 4*X + errors, X = X, e = errors)
Xh <- 10

fit_p2 <- lm(Y ~ X, data = dat)
b0_p2 <- fit_p2$coefficients[1]
b1_p2 <- fit_p2$coefficients[2]
Yh <- b0_p2 + b1_p2 * Xh

confint <- predict(fit_p2, data.frame(X=Xh), interval = "confidence", level = 0.95)

```

$b_0$ = `r round(b0_p2,4)`\
$b_1$ = `r round(b1_p2,4)`

$Y_h$ (at $X_h = 10$) = `r round(b0_p2,4)` + `r round(b1_p2,4)` \* 10 =
`r round(Yh,4)`

95% confidence interval: `r round(confint[2],3)` $\le E(Y_h) \le$
`r round(confint[3],3)`

b-)

```{r P2_b}
p2 <- replicate(200, {
  errors <- rnorm(n=5, mean=0, sd=5)
  dat_200 <- data.frame(Y = 20 + 4*X + errors, X = X, e = errors)
  fit_200 <- lm(Y ~ X, data = dat_200)
  b0_200 <- fit_200$coefficients[1]
  b1_200 <- fit_200$coefficients[2]
  conf <- predict(fit_200, data.frame(X=10), interval = "confidence", level = 0.95)
  
  out <- c(b0_200, b1_200, conf[1], conf[2], conf[3])
  out
})
p2 <- t(p2)
p2 <- data.frame(p2)
colnames(p2) <- c("b0","b1","Yh","lower","upper")
head(p2)
```

The results for 200 coefficients, $Y_h$, and lower/upper bounds of
confidence intervals are stored in the dataframe called 'p2', and its
first six rows are shown above.

c-)

```{r P2_c}
hist(p2$b1, breaks = 20, xlim = c(3.0, 5.5), main = "Frequency Distribution of the 200 Estimates of b1", xlab = expression(b[1]))

mean_b1 <- mean(p2[,2])
sd_b1 <- sqrt(var(p2[,2]))

# Theoretical Expectations
exp_b1mean <- 4 # coefficient of X in true regression
exp_b1sd <- sqrt(25 / sum((X - mean(X))^2)) # 25 is given in the problem as variance of E(Y)
```

Mean of 200 estimates of $b_1$ = `r round(mean_b1,4)`

SD of 200 estimates of $b_1$ = `r round(sd_b1,4)`

Theoretical expectation of mean of $b_1$ = `r exp_b1mean`

Theoretical expectation of SD of $b_1$ = `r round(exp_b1sd,4)`

For both mean and standard deviation, the estimates and the theoretical
expectations are consistent to each other as the numbers show above.

d-)

```{r P2_d}
mean_Yh <- mean(p2[,3])
notYh <- p2 %>% filter(lower > mean_Yh | upper < mean_Yh) %>% count()
propor <- notYh / nrow(p2)
```

`r propor*100`% of the 200 confidence intervals for $E(Y_h)$ when
$X_h = 10$ include $E(Y_h)$, and this result is consistent with
theoretical expectations as we had constructed 95% confidence intervals.

## Problem 3

```{r P3_setup}
question3 <- read.csv("/cloud/project/Data/Question3.csv")
```

a-)

```{r P3_a}
set.seed(1023)
train_index <- sort(sample(1:nrow(question3), 400))
train_set <- question3[train_index,]
test_set <- question3[-train_index,]
```

b-)

```{r P3_b}
N <- nrow(train_set)
fit_q3 <- lm(Y ~ X, data = train_set)
b0_q3 <- fit_q3$coefficients[1]
b1_q3 <- fit_q3$coefficients[2]
summary(fit_q3)
f2 <- qf(0.95,1,N-2)
```

Y = `r round(b0_q3,4)` + `r round(b1_q3,4)`X

Above regression model is significant as the summary shows that
F-statistic (922.5) is much bigger than $F(1-\alpha;1,n-2)$ =
`r round(f2,2)`, and p-value is very small.

c-)

```{r P3_c}
anova(fit_q3)
SSE.F <- anova(fit_q3)$`Sum Sq`[2]
f_q3 <- anova(fit_q3)$`Mean Sq`[1] / anova(fit_q3)$`Mean Sq`[2]
```

d-)

```{r P3_d}
fit_q3red <- lm(Y ~ 1, data = train_set)
b0_q3red <- fit_q3red$coefficients[1]
anova(fit_q3red)
SSE.R <- anova(fit_q3red)$`Sum Sq`[1]
```

Full Model: $Y_i = \beta_0 + \beta_1X_i + \epsilon_i =$
`r round(b0_q3,4)` + `r round(b1_q3,4)`$X_i$ + $\epsilon_i$

Reduced Model: $Y_i = \beta_0 + \epsilon_i =$ `r round(b0_q3red,4)` +
$\epsilon_i$

ANOVA tables for the full and reduced model show that SSE(R) =
`r SSE.R` drops significantly to SSE(F) =
`r SSE.F` when X is introduced to the full model. It
suggests that $H_a$ holds, and $F^* = \frac{MSR}{MSE}$
=`r round(f_q3,2)`\> $F(1-\alpha;1,n-2)$ =`r round(f2,2)`, so we
conclude $H_a$.

e-)

```{r P3_e}
par(mfrow=c(1,2))
y_hat <- fit_q3$fitted.values
resid_q3 <- fit_q3$residuals
plot(train_set$X, resid_q3, xlab = "X", ylab = "Residuals", main = "Residuals vs X")
plot(y_hat, resid_q3, xlab = expression(hat(Y)), ylab = "Residuals", main = "Residuals vs Y-hat")

par(mfrow=c(1,2))
mse <- anova(fit_q3)$`Mean Sq`[2]
semi_residual <- resid_q3 / sqrt(mse)
plot(train_set$X, semi_residual, xlab = "X", ylab = "Semistudentized Residuals", main = "Semistudentized Residuals vs X")
plot(y_hat, semi_residual, xlab = expression(hat(Y)), ylab = "Semistudentized Residuals", main = "Semistudentized Residuals vs Y-hat")
```

The normal residual plots show unequal variances as the dispersion of
data points gets wider as X or $\hat{Y}$ increase.

The semistudentized residual plots show that some points are about 4 or
more standard deviations away, which indicates that they could be
outliers.

f-)

```{r P3_f}
train_q3 <- data.frame(obs = train_set$Y, pred = fit_q3$fitted.values)
defaultSummary(train_q3)

predicted <- predict(fit_q3, test_set)
test_q3 <- data.frame(obs = test_set$Y, pred = predicted)
defaultSummary(test_q3)
```

g-)

```{r P3_g}
par(mfrow=c(1,2))
resid_test <- test_q3$obs - test_q3$pred 
plot(test_set$X, resid_test, xlab = "X", ylab = "Residuals", main = "Residuals vs X")
plot(test_q3$pred, resid_test, xlab = expression(hat(Y)), ylab = "Residuals", main = "Residuals vs Y-hat")

par(mfrow=c(1,2))
semi_residual2 <- resid_test / sqrt(mse)
plot(test_set$X, semi_residual2, xlab = "X", ylab = "Semistudentized Residuals", main = "Semistudentized Residuals vs X")
plot(test_q3$pred, semi_residual2, xlab = expression(hat(Y)), ylab = "Semistudentized Residuals", main = "Semistudentized Residuals vs Y-hat")
```

As it was shown the same in the training data, the test dataset also
show that the variances increase with X and $\hat{Y}$, and the
semistudentized residuals show that some data points are 4 standard
deviations away, indicating possible outliers.

h-)

Although we need to be careful, the model suggests that we should
consider removing outliers in the dataset. If we finally confirm the
existence of outliers after further analysis, we can remove them and
will be able to build a better model. The current model seems robust
because RMSE, R-Squared, and MAE values are consistent between training
and test datasets. Also, the R-Squared value slightly below 0.7
indicates that it is a fine model, but there is a possibility to improve
it by employing other predictor variables.
