---
title: "Assignment 5"
author: "Chul Hee Kim"
date: '2021 10 18 '
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("scipen" = 10)  ##Used to avoid displaying numbers such as 1+e3 (exponent), changing the options
# STEP 0: Function to load or download libraries
loadLib = function(libName)
{
  if(require(libName, character.only=TRUE))
  {
    ##cat(libName, "loaded properly\n")
  } else {
    cat("Installing", libName, "\n")
    install.packages(libName)
    if(require(libName, character.only=TRUE))
    {
      cat(libName, "loaded properly\n")
    } else {
      stop(c(libName, "not properly installed\n"))
    }
  }
}

libs = c("ggplot2","knitr","MASS","ggpubr","dplyr","tidyverse","faraway", 
         "rmarkdown", "markdown", "tidyr", "caret","readr","onewaytests","plyr",
         "lmtest","pander","kableExtra","QuantPsyc","GGally")

for (lib in libs)
{
  loadLib(lib)
}

## Helper to get data sets from one location on my R Cloud project
getDataSet <- function(dataset) {
  datasets.loc <- "/cloud/project/Data/"
  read.csv(paste0(datasets.loc,dataset,sep=""))
}
```

## Problem 1

a-)
```{r P1_a}
senic <- getDataSet("SENIC.csv")
X <- senic$AvailableFacilitiesPercent 
x <- scale(X, scale=FALSE)
xsq <- x^2
Y <- senic$Nurses
senic.reg <- lm(Y ~ x + xsq)
summary(senic.reg)
vif(senic.reg)
```
x = X - $\bar{X}$ \
Y = 150.079 + 7.066x + 0.101$x^2$

All the VIFs are almost 1, so there seems to be no collinearity problem.

b-)
```{r P1_b1, fig.height=4}
ei <- senic.reg$residuals
plot(senic.reg$fitted.values, ei, xlab="Fitted Values", ylab="Residuals")
```

The residuals against the fitted values plot shows possible nonconstant error variances and
potential outliers. We would probably need a further analysis (e.g., Box-Cox procedure), but
not shown here for the sake of this problem.

```{r P1_b2, fig.height=4}
senic %>% ggplot(aes(x=AvailableFacilitiesPercent, y=Nurses)) + 
  geom_point() + 
  stat_smooth(method='lm', formula=y~poly(x,2), se=FALSE) +
  xlab("AvailableFacilitiesPercent (scaled)")

```
The summary in part (a) shows $R^2$ = 0.6569, which is a fairly good number, and
the blue regression line above appears to be a good fit of the data points although
there is a higher variability for bigger X. The second-order regression model is
significant as well as both predictor variables, X and $X^2$.
\newpage
c-)
```{r P1_c}
senic.reduced <- lm(Y ~ x)
anova(senic.reduced, senic.reg)
f <- qf(0.9,1,110)
```
$H_0$: $\beta_2 = 0$ \
$H_a$: $\beta_2 \ne 0$

ANOVA table shows that $F^*$ = 13.806 > F(0.9;1,110) = `r round(f,2)`, so 
we conclude $H_a$ that the quadratic term cannot be dropped from the model.
p-value is 0.00032 and is less than $\alpha$=0.1.


## Problem 2

a-)
```{r P2_a}
data(divusa)
divusa2 <- divusa[,-1]
divusa.reg <- lm(divorce ~ ., data=divusa2)
summary(divusa.reg)
round(cor(divusa2),2)

xm <- model.matrix(divusa.reg)[,-1]
e <- eigen(t(xm)%*%xm)
k <- t(sqrt(e$val[1]/e$val))
kable(k, digits=3, align="c", caption="Condition Numbers", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
$R^2$ = 0.9208 is very high, and the p-value for the F-statistics is very small.
Hence, the model is significant, but two predictors 'unemplyed' and 'military'
could be insignificant depending on how we set the $\alpha$. There are some
large pairwise correlations, and the condition numbers shown in the table above
are all less than 30, so the collinearity does not appear to be an issue here.

b-)
```{r P2_b, message=FALSE}
vif(divusa.reg)
```
The VIFs shown above are all less than 10, which is the threshold level we are using here, 
so there is no evidence that collinearity causes issues here. But, one thing to keep
in mind is that 'femlab', 'marriage', and 'birth' are quite correlated to each other, 
so we may need a further analysis. For now, we move on to the next based solely on VIF results.

c-)
We use the significance level ($\alpha$=0.05), and the summary in part (a) shows that
'unemployed' and 'military' predictors are not significant as their p-values are greater
than 0.05.
```{r P2_c}
divusa.reduced <- lm(divorce ~ femlab+marriage+birth, data=divusa2)
summary(divusa.reduced)
anova(divusa.reduced, divusa.reg)
vif(divusa.reduced)
```
ANOVA table confirms that 'unemplyed' and 'military' can be dropped since the 
p-value for the F-statistic is greater than $\alpha$=0.05.
After removing the insignificant predictors from our model, all three VIFs are reduced,
so the collinearity seems to be reduced after the removal.


## Problem 3

a-)
```{r P3_a}
CP.data <- getDataSet("Commercial Properties Data Set.csv")
CP.reg <- lm(Y ~ ., data=CP.data)
lm.beta(CP.reg)
b1 <- lm.beta(CP.reg)[1]
b2 <- lm.beta(CP.reg)[2]
b3 <- lm.beta(CP.reg)[3]
b4 <- lm.beta(CP.reg)[4]
```
$\hat{Y}^*$ = `r round(b1,4)`$X_1^*$ + `r round(b2,4)`$X_2^*$ + `r round(b3,4)`$X_3^*$ + `r round(b4,4)`$X_4^*$

b-)
$b_1$: An increase of one standard deviation of $X_1$ when $X_2,X_3,X_4$ are fixed reduces Y by 0.548 sd. \
$b_2$: An increase of one standard deviation of $X_2$ when $X_1,X_3,X_4$ are fixed increases Y by 0.424 sd. \
$b_3$: An increase of one standard deviation of $X_3$ when $X_1,X_2,X_4$ are fixed increases Y by 0.048 sd. \
$b_4$: An increase of one standard deviation of $X_4$ when $X_1,X_2,X_3$ are fixed increases Y by 0.503 sd. \

We must be cautious about the relative importance of the standard regression coefficients because 
the magnitudes of the coefficients are affected not only by the presence of correlations among 
the predictor variables but also by the spacings of the observations on each of these variables.

c-)
```{r P3_c}
ybar <- mean(CP.data$Y)
x1bar <- mean(CP.data$X1)
x2bar <- mean(CP.data$X2)
x3bar <- mean(CP.data$X3)
x4bar <- mean(CP.data$X4)
s1 <- sd(CP.data$X1)
s2 <- sd(CP.data$X2)
s3 <- sd(CP.data$X3)
s4 <- sd(CP.data$X4)
sy <- sd(CP.data$Y)

ob1 <- (sy/s1)*b1
ob2 <- (sy/s2)*b2
ob3 <- (sy/s3)*b3
ob4 <- (sy/s4)*b4
ob0 <- ybar - ob1*x1bar - ob2*x2bar - ob3*x3bar - ob4*x4bar

tbl <- t(CP.reg$coefficients)

kable(tbl, digits=7, align="c", caption="Regression coefficients in the original variables", booktabs=TRUE) %>%
  kable_styling(latex_options="HOLD_position")
```
The regression model in the original variables (transformed back from the standardized regression): \
$\hat{Y}$ = `r round(ob0,3)`+`r round(ob1,3)`$X_1$+`r round(ob2,3)`$X_2$+`r round(ob3,3)`$X_3$+`r round(ob4,7)`$X_4$

The regression coefficients in the above equation are the same as the ones in the table, which are
the coefficients for the original regression model we obtained in Problem 4b in HW4.

d-)
```{r P3_d}
vif(CP.reg)
mean.vif <- mean(vif(CP.reg))
```
The maximum of the VIFs is 1.648 and the mean of them is `r round(mean.vif,3)`. 
It means that the expected sum of squared errors in the least squares standardized
regression coefficients is about 1.4 times as large as it would be if the X variables
were uncorrelated. Therefore, X variables are close to being uncorrelated. Also,
all four VIFs are significantly less than 10 and even close to 1, so 
we do not see much multicollinearity problem here.
