---
title: "Assignment 8"
author: "Chul Hee Kim"
date: '2021 11 5 '
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("scipen" = 10)  ##Used to avoid displaying numbers such as 1+e3 (exponent), changing the options
# STEP 0: Function to load or download libraries
loadLib = function(libName)
{
  if(require(libName, character.only=TRUE))
  {
    ##cat(libName, "loaded properly\n")
  } else {
    cat("Installing", libName, "\n")
    install.packages(libName)
    if(require(libName, character.only=TRUE))
    {
      cat(libName, "loaded properly\n")
    } else {
      stop(c(libName, "not properly installed\n"))
    }
  }
}

libs = c("ggplot2","knitr","MASS","ggpubr","dplyr","tidyverse","faraway", 
         "rmarkdown", "markdown", "tidyr", "caret","readr","onewaytests","plyr",
         "lmtest","pander","kableExtra","QuantPsyc","GGally","olsrr","leaps",
         "ggcorrplot","glmnet","rpart","rpart.plot","C50","gmodels","visdat")

for (lib in libs)
{
  loadLib(lib)
}

## Helper to get data sets from one location on my R Cloud project
getDataSet <- function(dataset) {
  datasets.loc <- "/cloud/project/Data/"
  read.csv(paste0(datasets.loc,dataset,sep=""))
}
```

## Problem 1

a-)
```{r P1_a}
data(diabetes)
x <- diabetes[,-6]
par(mfrow=c(2,2))
for (i in 1:18){
  plot(x[,i], diabetes$glyhb, xlab=colnames(x)[i], ylab="glyhb")
}
```

Among the 18 predictors, it seems that only 'stab.glu' has the apparent positive 
relationship with the response 'glyhb'. We also observe that 'age' or 'frame'
might have significant relationship, but it is not very clear in the plots.

b-)
```{r fig.height=3}
vis_miss(diabetes)
```

We see that there are a bunch of NAs for 'bp.2s' and 'bp.2d',
so we remove these two predictors since removing them by rows will result in 
262 fewer observations, which is 65% of the total observations. Next, we find 
that there are some remaining NAs, which can be eliminated by \emph{na.omit} function.
```{r P1_b}
diabetes2 <- diabetes[,-c(15,16)]
diabetes2 <- na.omit(diabetes2)
table(is.na(diabetes2))
```
We finally see that there is no more NA in the data.

c-)
```{r P1_c}
f1 <- lm(glyhb ~ ., data=diabetes2)
summary(f1)
ols_step_both_p(f1, prem=0.05)
plot(f1$fitted.values, f1$residuals, xlab="Fitted Values", ylab="Residuals")
```
'stab.glu', 'chol', 'age', 'ratio', and 'time.ppn' appears to be the most important predictors
as suggested by the stepwise selection summary above.
'stab.glu', 'age', and 'time.ppn' are also indicated as the important predictors by
the linear regression summary. 'stab.glu' appears to be the most important of all.

The residuals against fitted values plot shows that the error variances
are fairly constant with a few potential outliers. One might argue that
the variability increases with larger fitted values, but we will leave it
as is for now and move on to the next.

d-)
```{r P1_d}
f.rt <- rpart(glyhb ~ ., data=diabetes2)
f.rt
median(diabetes2$glyhb)
```
For 'stab.glu < 158', there are 324 observations, and the mean response for them
is 5.028333. The terminal node is indicated with * on the right-hand side in the
results shown above, and the node#8 'age < 55.5' is the largest with 205 observations
and the mean response of 4.598683, which is close to the median of 'glyhb' (4.86).

e-)
```{r P1_e}
rpart.plot(f.rt, digits = 3, fallen.leaves = TRUE, type = 3, extra = 101)
```

The predictor that splits the tree at the top is the most important predictor.
In our case, 'stab.glu' is the most important, and it splits the data into two
categories 'stab.glu < 158' and 'stab.glu >= 158'.

f-)
```{r P1_f}
yhat.rt <- predict(f.rt, diabetes2)
ei <- diabetes2$glyhb - yhat.rt
plot(yhat.rt, ei, xlab="Fitted Values", ylab="Residuals")

MAE <- function(actual, predicted) {mean(abs(actual - predicted))}
mae1 <- MAE(diabetes2$glyhb, yhat.rt)
mae2 <- MAE(diabetes2$glyhb, f1$fitted.values)

SSE <- function(actual, predicted) {sum((actual - predicted)^2)}
sse1 <- SSE(diabetes2$glyhb, yhat.rt)
sse2 <- SSE(diabetes2$glyhb, f1$fitted.values)

R2 <- function(actual, predicted) {sum((actual - predicted)^2)/((length(actual)-1)*var(actual))}
rsq1 <- 1-R2(diabetes2$glyhb, yhat.rt)
rsq2 <- 1-R2(diabetes2$glyhb, f1$fitted.values)

tbl1 <- data.frame(rbind(c(mae1,mae2), c(sse1,sse2), c(rsq1,rsq2)))
colnames(tbl1) <- c("Regresstion Tree", "Linear Regression")
rownames(tbl1) <- c("MAE", "SSE", "Rsquared")
kable(tbl1, digits=3, align="c", caption="Regression Tree vs. Linear Regression", 
      booktabs=TRUE) %>% kable_styling(latex_options="HOLD_position")
```

The residuals against the fitted values plot shows the constant error variances and
looks a little bit better than the one for the linear regression model.
We would choose the regression tree model over the linear regression model because
as shown in the table above, MAE and SSE is lower and $R^2$ is higher for the 
regression tree model.


## Problem 2

a-)
```{r P2_a}
mush <- getDataSet("Mushroom.csv")
mush$Class <- as.factor(mush$Class)
mush$veil.type <- NULL # removed because all the observations are the same.
mush$stalk.root <- NULL # removed because of too many observations marked as '?'
# Tried including 'stalk.root' and no difference in the result
set.seed(1023)
train.index <- createDataPartition(y = mush$Class, p = 0.9, list = FALSE)
train.set <- mush[train.index,]
test.set <- mush[-train.index,]
```


b-)
```{r P2_b}
tree.model <- C5.0(train.set[,-1], train.set[,1])
tree.model
summary(tree.model)
```
The decision tree shown above clearly indicates which sub-tree leads to
'p'(poisonous) and 'e'(edible).

c-)
```{r P2_c}
pred.train <- predict(tree.model, train.set)
CrossTable(train.set$Class, pred.train, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))
```
The error rate is 0 since the table above shows that the model correctly predicts
for all the observations of both 'e' and 'p'.

d-)
```{r P2_d}
pred.test <- predict(tree.model, test.set)
CrossTable(test.set$Class, pred.test, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))
```
The model also correctly predicts all the observations in the test data, so the model
is very stable. 


## Problem 3

a-)
```{r P3_a}
x <- model.matrix(glyhb~., diabetes2)[,-c(1)]
y <- diabetes2$glyhb
RidgeMod <- glmnet(x, y, alpha=0, nlambda=100,lambda.min.ratio=0.0001)
CvRidgeMod <- cv.glmnet(x, y, alpha=0, nlambda=100,lambda.min.ratio=0.0001)
plot(CvRidgeMod)
best.lambda.ridge <- CvRidgeMod$lambda.min
pred.ridge <- coef(CvRidgeMod, s = "lambda.min")
pred.ridge
```

b-)
```{r P3_b}
LassoMod <- glmnet(x, y, alpha=1, nlambda=100,lambda.min.ratio=0.0001)
CvLassoMod <- cv.glmnet(x, y, alpha=1, nlambda=100,lambda.min.ratio=0.0001)
plot(CvLassoMod)
best.lambda.lasso <- CvLassoMod$lambda.min
pred.lasso <- coef(CvLassoMod, s = "lambda.min")
pred.lasso
```
c-)
```{r P3_c}
EnetMod <- glmnet(x, y, alpha=0.5, nlambda=100,lambda.min.ratio=0.0001)
CvElasticnetMod <- cv.glmnet(x, y,alpha=0.5,nlambda=100,lambda.min.ratio=0.0001)
best.lambda.enet <- CvElasticnetMod$lambda.min
pred.enet <- coef(CvElasticnetMod, s = "lambda.min")
pred.enet
```

d-)
```{r P3_d}
rr.huber <- rlm(glyhb ~ ., data = diabetes2, psi = psi.huber)
summary(rr.huber)

rr.bisquare <- rlm(glyhb ~ ., data = diabetes2, psi = psi.bisquare)
summary(rr.bisquare)
```

e-)
```{r P3_e}
yhat.ridge <- predict(RidgeMod, s = best.lambda.ridge, newx = x)
yhat.lasso <- predict(LassoMod, s = best.lambda.lasso, newx = x)
yhat.enet <- predict(EnetMod, s = best.lambda.enet, newx = x)
yhat.huber <- predict(rr.huber, diabetes2)
yhat.bisquare <- predict(rr.bisquare, diabetes2)

mae.ridge <- MAE(diabetes2$glyhb, yhat.ridge)
mae.lasso <- MAE(diabetes2$glyhb, yhat.lasso)
mae.enet <- MAE(diabetes2$glyhb, yhat.enet)
mae.huber <- MAE(diabetes2$glyhb, yhat.huber)
mae.bisquare <- MAE(diabetes2$glyhb, yhat.bisquare)

sse.ridge <- SSE(diabetes2$glyhb, yhat.ridge)
sse.lasso <- SSE(diabetes2$glyhb, yhat.lasso)
sse.enet <- SSE(diabetes2$glyhb, yhat.enet)
sse.huber <- SSE(diabetes2$glyhb, yhat.huber)
sse.bisquare <- SSE(diabetes2$glyhb, yhat.bisquare)

rsq.ridge <- 1-R2(diabetes2$glyhb, yhat.ridge)
rsq.lasso <- 1-R2(diabetes2$glyhb, yhat.lasso)
rsq.enet <- 1-R2(diabetes2$glyhb, yhat.enet)
rsq.huber <- 1-R2(diabetes2$glyhb, yhat.huber)
rsq.bisquare <- 1-R2(diabetes2$glyhb, yhat.bisquare)

ridge.pf <- c(mae.ridge, sse.ridge, rsq.ridge)
lasso.pf <- c(mae.lasso, sse.lasso, rsq.lasso)
enet.pf <- c(mae.enet, sse.enet, rsq.enet)
huber.pf <- c(mae.huber, sse.huber, rsq.huber)
bisquare.pf <- c(mae.bisquare, sse.bisquare, rsq.bisquare)

tbl2 <- cbind(tbl1, ridge.pf, lasso.pf, enet.pf, huber.pf, bisquare.pf)
colnames(tbl2) <- c("Regression Tree","OLS","Ridge","Lasso","ElasticNet","Huber","Bisquare")
kable(tbl2, digits=3, align="c", caption="Comparison of Model Performance", 
      booktabs=TRUE) %>% kable_styling(latex_options="HOLD_position")
```
The regression tree performs the best, so we would choose it as the final model. 