---
title: "Assignment 9"
author: "Chul Hee Kim"
date: '2021 11 12 '
output: pdf_document
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(ECHO=TRUE)
options("scipen" = 10)

loadLib = function(libName)
{
    if(require(libName, character.only=TRUE))
    {
        ##cat(libName, "loaded properly\n")
    } else {
        cat("Installing", libName, "\n")
        install.packages(libName)
        if(require(libName, character.only=TRUE))
        {
            cat(libName, "loaded properly\n")
        } else {
            stop(c(libName, "not properly installed\n"))
        }
    }
}

libs = c("ggplot2", "knitr", "MASS", "car", "onewaytests", "olsrr","ggpubr","dplyr", "Hmisc", "faraway", 
         "rmarkdown", "markdown", "tidyr", "caret","kableExtra", "rpart", "rpart.plot", "lmtest", "Metrics", "GGally",
         "adabag", "vcd", "randomForest","ipred","kernlab","vcd","gbm","C50","gmodels")

for (lib in libs)
{
    loadLib(lib)
}

## Helper to get data sets from one location on hard drive
getDataSet <- function(dataset) {
  datasets.loc <- "/cloud/project/"
  read.csv(paste0(datasets.loc,dataset,sep=""))
}
```

## Problem 1

a-)
```{r P1_a1}
data("uswages")
head(uswages)
str(uswages)

uswages.sub <- uswages[,6:9]
table(rowSums(uswages.sub)==1)
```
There are four categorical variables related to region, and we want to check if
all of them are necessary. The result above tells us that the sum of every row
of \emph{ne, mw, so, we} predictors is 1, which means that one of four variables
should be removed. We will remove \emph{we} from \emph{uswages} dataset.

```{r}
uswages <- uswages[,-9] # remove 'we' variable

f1 <- lm(wage ~ ., data=uswages)
summary(f1)
```
The initial regression model including all predictors shows insignificant variables
if $\alpha$=0.05, but it can be argued since their p-values are very close to 0.05. 
$R^2$ is 0.2, which is not very good.

```{r P1_a2}
ols_step_both_p(f1, prem=0.05)
k1 <- ols_step_best_subset(f1)
k1
plot(k1)
```

The stepwise regression selects five predictors, \emph{educ, exper, pt, smsa, race}.
The best subset regression also selects the same five predictors since Model 5 
shows the best results given by adjusted $R^2$, $C_p$, and AIC. Now, let's check
the regression model assumptions.

```{r P1_a3}
f1a <- lm(wage ~ educ+exper+pt+smsa+race, data=uswages)
summary(f1a)
par(mfrow=c(2,2))
plot(f1a)

vif(f1a)

dwtest(f1a)
```

The residuals against fitted values plot shows a slight cone shape, and the 
scale-location plot shows $\sqrt{standardized residuals}$ increasing as the
fitted values becomes larger. The QQ plot indicates that the error terms are
approximately normally distributed but a little skewed right. All these plots
indicate a few potential outliers.

VIFs show no multicollinearity issue.

Durbin-Watson test concludes that there is autocorrelation in the model.

b-)
```{r P1_b}
f1b <- lm(log(wage) ~ educ+exper+pt+smsa+race, data=uswages)
summary(f1b)

ols_step_both_p(f1b, prem=0.05)

f1b2 <- lm(log(wage) ~ educ+exper+pt+smsa+race, data=uswages)
summary(f1b2)
par(mfrow=c(2,2))
plot(f1b2)

vif(f1b2)

dwtest(f1b2)
```
The residuals against fitted values plot no longer shows the cone shape, and the 
scale-location plot shows the red horizontal line, which is improved from the
previous model. The QQ plot indicates that the error terms are
approximately normally distributed but still a little skewed right. All these plots
still indicate a few potential outliers.

VIFs show no multicollinearity issue.

Durbin-Watson test concludes that there is now no autocorrelation in the model.

$R^2$ is improved from 0.2 in part(a) to 0.38 in part(b), 
so all the results mentioned above indicate that the log-transformed
model is the one we would choose over the one in part(a).

c-)
```{r P1_c}
f1tree <- rpart(wage ~ ., data=uswages)
f1tree

pred.tree <- predict(f1tree, uswages)
summary(pred.tree)

MAE <- function(actual, predicted) {mean(abs(actual - predicted))}
SSE <- function(actual, predicted) {sum((actual - predicted)^2)}
R2 <- function(actual, predicted) {sum((actual - predicted)^2) /((length(actual)-1)*var(actual))}

mae.c <- MAE(uswages$wage, pred.tree)
sse.c <- SSE(uswages$wage, pred.tree)
r2.c <- 1-R2(uswages$wage, pred.tree)
pred.c <- c(mae.c, sse.c, r2.c)
```

d-)
```{r P1_d}
f1tree.log <- rpart(log(wage) ~ ., data=uswages)
f1tree.log

pred.tree.log <- predict(f1tree.log, uswages)
summary(pred.tree.log)

mae.d <- MAE(uswages$wage, exp(pred.tree.log))
sse.d <- SSE(uswages$wage, exp(pred.tree.log))
r2.d <- 1-R2(uswages$wage, exp(pred.tree.log))
pred.d <- c(mae.d, sse.d, r2.d)

tbl <- cbind(pred.c, pred.d)
colnames(tbl) <- c("Normal Tree", "Log-scaled Tree")
rownames(tbl) <- c("MAE", "SSE", "Rsquared")
kable(tbl, digits=3, align="c", caption="Comparison of Model Performance", 
      booktabs=TRUE) %>% kable_styling(latex_options="HOLD_position")
```

The results are mixed as MAE is smaller for the log-scaled tree model, 
$R^2$ is higher for the normal tree model,
but the differences are very small. There is no definite winner here, but
given similar performances, we would choose the normal tree model 
because it is the simpler model and is easier to interpret. 

e-)
```{r P1_e}
set.seed(1023)
rf <- randomForest(wage ~ ., data = uswages)
rf
pred.rf <- predict(rf, uswages)
mae.e <- MAE(uswages$wage, pred.rf)
sse.e <- SSE(uswages$wage, pred.rf)
r2.e <- 1-R2(uswages$wage, pred.rf)
pred.e <- c(mae.e, sse.e, r2.e)
```

f-)
```{r P1_f}
set.seed(1023)
bg <- ipred::bagging(wage ~ ., data = uswages, nbagg = 25)
pred.bg <- predict(bg, uswages)
mae.f <- MAE(uswages$wage, pred.bg)
sse.f <- SSE(uswages$wage, pred.bg)
r2.f <- 1-R2(uswages$wage, pred.bg)
pred.f <- c(mae.f, sse.f, r2.f)
```

g-)
```{r P1_g}
set.seed(1023)
bt <- gbm(wage ~ ., data = uswages)
pred.bt <- predict(bt, uswages)
mae.g <- MAE(uswages$wage, pred.bt)
sse.g <- SSE(uswages$wage, pred.bt)
r2.g <- 1-R2(uswages$wage, pred.bt)
pred.g <- c(mae.g, sse.g, r2.g)

# Comparison of all models
tbl2 <- cbind(pred.c, pred.d, pred.e, pred.f, pred.g)
colnames(tbl2) <- c("Normal Tree", "Log-scaled Tree", "Random Forest", "Bagging", "Boosting")
rownames(tbl2) <- c("MAE", "SSE", "Rsquared")
kable(tbl2, digits=3, align="c", caption="Comparison of Model Performance", 
      booktabs=TRUE) %>% kable_styling(latex_options="HOLD_position")
```
The table above suggests that the random forest performs significantly better than all other models.


## Problem 2

a-)
```{r P2_a}
data("pima")
head(pima)
str(pima)

pima$test <- as.factor(pima$test)
set.seed(1023)
train.index <- createDataPartition(y = pima$test, p = 0.9, list = FALSE)
train.set <- pima[train.index,]
test.set <- pima[-train.index,]

f2t <- C5.0(train.set[,-9], train.set[,9])
f2t
summary(f2t)
pred.f2t <- predict(f2t, train.set)
CrossTable(train.set$test, pred.f2t, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))
```

b-)
```{r P2_b}
f2.svm <- ksvm(test ~ ., data = train.set, kernel = "vanilladot")
f2.svm2 <- ksvm(test ~ ., data = train.set, kernel = "rbfdot")

pred.svm <- predict(f2.svm, train.set)
pred.svm2 <- predict(f2.svm2, train.set)

CrossTable(train.set$test, pred.svm, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))
CrossTable(train.set$test, pred.svm2, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))
```
We chose two of the most common kernels for our SVM, linear ("vanilladot") and Gaussian RBF kernel ("rbfdot").
The error rate is 160/692 = 23.12% for the linear kernel and 129/692 = 18.64% for the Gaussian RBF kernel.
It seems that the latter performs better.

c-)
```{r P2_c}
pred.f2t.test <- predict(f2t, test.set)
pred.svm.test <- predict(f2.svm, test.set)
pred.svm2.test <- predict(f2.svm2, test.set)

CrossTable(test.set$test, pred.f2t.test, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))
CrossTable(test.set$test, pred.svm.test, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))
CrossTable(test.set$test, pred.svm2.test, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default'))
```
The error rate is 16/76 = 21.05% for the tree model, 12/76 = 15.79% for SVM with the linear kernel, 11/76 = 14.47% 
for SVM with the Gaussian RBF kernel. The difference between the two SVM models is marginal, but
the RBF kernel model also performed better with the training data, so we would select the SVM 
with the Gaussian RBF kernel as our final model.