---
title: "Assignment 10"
author: "Chul Hee Kim"
date: '2021 11 23 '
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(ECHO=TRUE)

# STEP 1: Function to load or download libraries
loadLib = function(libName)
{
    if(require(libName, character.only=TRUE))
    {
        ##cat(libName, "loaded properly\n")
    } else {
        cat("Installing", libName, "\n")
        install.packages(libName)
        if(require(libName, character.only=TRUE))
        {
            cat(libName, "loaded properly\n")
        } else {
            stop(c(libName, "not properly installed\n"))
        }
    }
}

libs = c("ggplot2", "knitr", "MASS", "car", "onewaytests", "olsrr","ggpubr","dplyr", "Hmisc", "faraway","forecast", 
         "rmarkdown", "markdown", "tidyr", "caret","kableExtra", "rpart", "rpart.plot", "lmtest", "Metrics","kernlab",      
         "GGally","visdat","e1071","adabag", "vcd", "randomForest","ggcorrplot","mboost","party","neuralnet",
         "HoRM","Hmisc","orcutt")

for (lib in libs)
{
    loadLib(lib)
}

## Helper to get data sets from one location on hard drive
getDataSet <- function(dataset) {
  datasets.loc <- "/cloud/project/Data/"
  read.csv(paste0(datasets.loc,dataset,sep=""))
}
## Helper to display regression function with n coefficients
## If var.names = TRUE add a space and "*" symbol to make readable
dispRegFunc <- function(reg, var.names=FALSE, pdfWrap=0, signif=6) {

  ## Set the line wrapping only if this is being knit as a pdf
  line.limit <- 0
  if (!is.null(knitr::current_input())) {
    if("pdf_document" %in% rmarkdown::all_output_formats(knitr::current_input())) {
       line.limit <- pdfWrap
    }
  }
  ## Check to see if this is a cv.glmnet object or a lm or rlm object
  ifelse(class(reg) == "cv.glmnet",
         coefs <- predict(reg, s=reg$lambda.min, type="coefficients")[,1],
         coefs <- reg$coefficients)
  
  b0 = coefs[1]
  n <- length(coefs)
  my_formula <- paste0("Y = ", round(b0, digits = signif))
  for (i in 2:n) {
    ## Choose the +/- sign for the coefficient then remove it using abs()
    new.formula.term <- paste0(ifelse(coefs[i] < 0," - "," + "),
                         signif(abs(coefs[i]),signif),
                         ifelse(var.names,paste0(" * ",names(coefs)[i]),
                         names(coefs)[i]))
    if (line.limit > 0) {
      if (line.limit <= nchar(my_formula) + nchar(new.formula.term)) {
        print(my_formula)
        my_formula <- new.formula.term
      } else {
        my_formula <- paste0(my_formula, new.formula.term)
      }
    }
    else {
      my_formula <- paste0(my_formula, new.formula.term)
    }
  }
  print(my_formula)
}

## Helper function to calculate R2, RMSE, and MAE using the 1-SSE/SST method of R2
## instead of the postResample() method of squaring the correlation coefficient.
resampleMetrics <- function(pred, obs) {
	t(c(R2=caret::R2(pred, obs, formula = "traditional"),postResample(pred,obs)[-2]))
}
set.seed(123)
options("scipen" = 10)  ##Used to avoid displaying numbers such as 1+e3 (exponent), changing the options

## Function to calculate the optimal Cp for pruning
getBestCp <- function(full_tree) {
  ## Get the cp (complexity parameter) table for the full tree
  cptable <- full_tree$cptable
  nsplit <- nrow(cptable)
  ## Get the minimum value of xerror (we want to choose the xerror within 1 STD 
  ## of the minimum xerror with the the lowest nsplit value)
  min.xerror <- which.min(full_tree$cptable[, 4]) 
  for(i in 1:nsplit) {
    ## Return the first value within 1 STD of the minimum xerror,
    ## which will have the lowest nsplit due loop order
    if(cptable[i, 4] < cptable[min.xerror, 4] + cptable[min.xerror, 5]) return(cptable[i, 1])
  }
  ## Didn't find anything, return the full tree
  return(cptable[nsplit, 1])
}
```

## Problem 1

a-)
Out of 97 total observations, two predictors, \emph{benign.prostatic.hyperplasia} and 
\emph{capsular.penetration} have too many zeros. We need to first examine what these mean.
Roughly 45% of 97 observations are zero for the two variables, so we may be tempted 
to completely ignore these two. But, a little bit of research showed that having 0
for those two metrics is common, especially for healthy men. Therefore, we will
leave them as is in our dataset and start analyzing. The correlation matrix shows 
a few important variables, and we will use the best subset and stepwise both regression 
to find our model.
```{r P1_a}
prostate <- getDataSet("Prostate Cancer.csv")
str(prostate)
head(prostate)
apply(prostate,2,function(x) sum(is.na(x))) # check any missing data
apply(prostate,2,function(x) sum(x==0)) # check zeros

corr <- round(cor(prostate),2)
ggcorrplot(corr, type="lower", lab=TRUE)

set.seed(567)
train.index <- sample(nrow(prostate), 65)
train.set <- prostate[train.index,]
test.set <- prostate[-train.index,]

f1 <- lm(PSA.level ~ ., data = train.set)
summary(f1)

k1 <- ols_step_best_subset(f1)
k1
plot(k1)

ols_step_both_p(f1, prem=0.05)
```

The best subset regression shows that the choice could be Model 4 with \emph{Cancer.volume, Weight, Age, Capsular.penetration}. But, using these four predictors results in a negative $R^2$ when assessed with
the test set. The reason is that the OLS regression function has a negative beta for \emph{Age}, so 
when the model is tested on the test set, many predicted PSA levels became negative and totally different
from the observed values. We go back to the best subset regression summary and see that Model 4 has 
only a marginal advantage over Model 2 and 3. Our final choice is Model 2 because it is simpler
and performs just as well. The stepwise both regression also confirms our choice.

```{r P1_a2}
m2 <- lm(PSA.level~Cancer.volume+Capsular.penetration, data=train.set)
summary(m2)
dispRegFunc(m2)
resampleMetrics(predict(m2, test.set), test.set$PSA.level)
```

We should check the regression model assumptions and outliers before we talk
about the model's performance and validity.

```{r fig.height=4}
f1.1 <- lm(PSA.level ~ Cancer.volume+Capsular.penetration, data = prostate)
n <- nrow(prostate)
p <- length(f1.1$coefficients)

outlierTest(f1.1)
ols_plot_resid_stud_fit(f1.1)
ols_plot_resid_stud(f1.1)
cook <- cooks.distance(f1.1)
plot(cook, ylab="Cooks distance")
ols_plot_dffits(f1.1)
hatv <- hatvalues(f1.1)
plot(hatv, ylab="Leverages",main="Leverages vs Index Plot")
abline(h=2*p/n)
plot(f1.1, which=5, col=c("blue"))
```
The points 95,96,97 are identified as outliers or at least as very influential points.
But given a small dataset, it is hard to justify the removal of these points, so
we will keep them in our dataset and continue.

```{r P1_a3}
par(mfrow=c(2,2))
plot(m2)
vif(m2)
shapiro.test(m2$residuals)
```

The residual vs fitted plot and the scale-location plot indicate nonconstant error variances.
The QQ plot and Shapiro-Wilk test show non-normality. There is no multicollinearity issue.
Let us try the transformation and see if we can remediate these.

```{r}
f.full <- lm(PSA.level ~ Cancer.volume+Capsular.penetration, data=prostate)
bc <- boxcox(f.full,seq(-1,2,0.01))
best.lambda <-bc$x[which(bc$y==max(bc$y))]
best.lambda

f1.log <- lm(log(PSA.level) ~ Cancer.volume+Capsular.penetration, data = train.set)
summary(f1.log)
par(mfrow=c(2,2))
plot(f1.log)
vif(f1.log)
shapiro.test(f1.log$residuals)

p.ols <- resampleMetrics(exp(predict(f1.log, test.set)), test.set$PSA.level)
p.ols
```
The best $\lambda$ for Box-Cox procedure is 0.11, but since $Y^{0.11}$ transformation
is complicated, we choose the log transformation, which seem to remediate the problems
we had. Both constant error variance and normality assumption now hold, and 
there is no multicollinearity. However, now the performance of the model is poor
with $R^2$ of less than 10%. With this poor performance, it would not be useful for
the oncologists to predict PSA levels. We could try Ridge, Lasso, or robust regression
to mitigate the influence of outliers, but for the sake of this problem, we settle
with the best subset model and move on.

b-)
```{r P1_b}
f.tree <- rpart(PSA.level ~ ., data = train.set)
f.tree
f.tree$cptable
plotcp(f.tree)
rpart.plot(f.tree, digits=3)
p.tree <- resampleMetrics(predict(f.tree,test.set), test.set$PSA.level)
```
Although the $C_p$ table suggests that we prune the tree since the entire tree falls
into the minimum 'xerror'+ 1 SE, but by this standard, we prune too much and have
no tree at all. Therefore, we keep the original tree as is, and it is already a
simple tree.

The regression tree only considers \emph{Cancer.volume}, and if it is greater than
or equal to 16.53, PSA level is predicted to be 91.45. When it is less than 4.71,
PSA level = 8.23, and when greater than or equal to 4.71 and less than 16.53,
PSA level = 20.82.

c-)
```{r P1_c}
table1 <- rbind(p.ols, p.tree)
rownames(table1) <- c("OLS","Regression Tree")
kable(table1, digits=3, align="c", booktabs=TRUE,caption="Model Comparison") %>%
  kable_styling(latex_options="HOLD_position")
```

The regression tree performs better and is easier to interpret since it only takes
into account one variable and also has the plot of the tree. Also, the best regression 
model is log transformed, which makes it harder to interpret, and interpreting 
the regression function is more complicated at least for this problem, given such
a simple regression tree.

d-)
Neural networks work best when the input data are scaled to a narrow range around zero,
so we first normalize each column of the dataset. The normalization also leads to
faster learning and convergence.

```{r P1_d}
normalize <- function(x) {return((x - min(x)) / (max(x) - min(x)))}
prostate_norm <- as.data.frame(lapply(prostate, normalize))

train_norm <- prostate_norm[train.index,]
test_norm <- prostate_norm[-train.index,]

softplus <- function(x) { log(1 + exp(x)) }
set.seed(567)
NN <- neuralnet(PSA.level ~ ., data = train_norm, hidden = 4, act.fct = softplus)
plot(NN)
NN_result <- compute(NN, test_norm[2:8])
pred.NN <- NN_result$net.result
cor(pred.NN, test_norm$PSA.level)

unnormalize <- function(x) {return((x * max(prostate$PSA.level) - min(prostate$PSA.level)) + min(prostate$PSA.level))}
p.NN <- resampleMetrics(unnormalize(pred.NN), test.set$PSA.level)

# check in-sample performance to see if there's overfitting
NN_result2 <- compute(NN, train_norm[2:8])
pred.NN2 <- NN_result2$net.result
cor(pred.NN2, train_norm$PSA.level)

p.NN2 <- resampleMetrics(unnormalize(pred.NN2), train.set$PSA.level)
```

ReLU is the most popular activation function, and it is what we also want to use,
but since it doesn't work with neuralnet() function, we will instead use a smooth
approximation of ReLU, i.e. softplus. 

For the choice of number of hidden layers and nodes, we first choose just one
hidden layer, given our small dataset. Four hidden nodes are chosen since it is
customary to choose between the number of input nodes and that of output nodes, 
which are 7 and 1 in this case. 

The interpretation of a neural network is very difficult, but based on what we
see in the plot above, each of the 7 input nodes is sent to each of the hidden nodes
with different weights and biases. Each hidden node converts these inputs with
the activation function, which is softplus in our case, and again with different
weights and bias, they are sent to the single output node, \emph{PSA.level}.

The problem with our model is that it significantly overfits the training data.
We are already using a small number of hidden nodes, and using only one hidden node
does not improve the overfitting. We probably need a lot more data to effectively
use the neural network, so for this problem, there appears not much for us to do.

```{r}
table1.1 <- rbind(p.NN, p.NN2)
rownames(table1.1) <- c("Test Data","Training Data")
kable(table1.1, digits=3, align="c", booktabs=TRUE,caption="Overfitting") %>%
  kable_styling(latex_options="HOLD_position")
```

e-)
```{r P1_e}
table2 <- rbind(p.ols, p.tree, p.NN)
rownames(table2) <- c("OLS","Regression Tree","Neural Network")
kable(table2, digits=3, align="c", booktabs=TRUE,caption="Model Comparison") %>%
  kable_styling(latex_options="HOLD_position")

```
The regression tree is our best model as shown in the table above.

f-)
In this problem, we covered three models, ordinary least squares regression,
regression tree, and neural network. Of these three, the neural network is the
most difficult one to interpret since by nature, this model uses all variables 
with many weights and biases involved along with the activation function. Between
the OLS and the regression tree, which one is easier depends on how complicated 
the regression tree happens to be. Once the regression tree includes too many
splits it gets harder and harder to interpret while OLS just depends on how many
variables it uses. As previously mentioned, in this problem, the regression tree
is very simple and is easier to interpret.


## Problem 2

a-)
```{r P2_a}
AA <- getDataSet("Advertising Agency.csv")
str(AA)
head(AA)

f2 <- lm(Y ~ X, data=AA)
summary(f2)
et <- f2$residuals
dwtest(Y~X, data=AA)
```

$H_o$: $\rho$ = 0 \
$H_a$: $\rho$ > 0 

The Durbin-Watson test shows that the p-value is less than $\alpha$=0.01, so
we reject the null hypothesis. There seems to exist the autocorrelation.

b-)
```{r P2_b}
et1 <- Lag(et, shift = 1)

d1 <- sum(na.omit(et1*et))
d2 <- sum(na.omit(et1^2))
rho <- d1/d2

Ytnew <- AA$Y - rho * Lag(AA$Y, shift = 1)
Xtnew <- AA$X - rho * Lag(AA$X, shift = 1)

f2.1 <- lm(Ytnew ~ Xtnew)
summary(f2.1)
b0p <- f2.1$coefficients[1]
b1p <- f2.1$coefficients[2]
sb0p <- summary(f2.1)$coefficients[1,2]
sb1p <- summary(f2.1)$coefficients[2,2]
dwtest(Ytnew ~ Xtnew)
```
The Durbin-Watson test result suggest that we fail to reject the null hypothesis, so
after the first iteration of the Cochrane-Orcutt procedure, the autocorrelation
seems to be disappeared.

c-)
```{r P2_c}
b0 <- b0p / (1-rho)
b1 <- b1p
sb0 <- sb0p / (1-rho)
sb1 <- sb1p
```
After the first iteration of the Cochrane-Orcutt:\
$\hat{Y'}$ = `r round(b0p,3)` + `r round(b1p,3)`X' \

In terms of the original variables:\
$\hat{Y}$ = `r round(b0,3)` + `r round(b1,3)`X

The coefficients for X and X' are the same, and those for the intercepts are different.

$s(b_0)$ = `r round(sb0,4)` \
$s(b_1)$ = `r round(sb1,4)`

d-)
```{r P2_d}
MSE <- summary(f2.1)$sigma^2
Xp <- Xtnew
Xbar.p <- mean(Xp[-1])

X20 <- AA$X[20]
X21 <- 3.625
X21p <- X21 - rho * X20

Yhat21 <- b0 + b1 * X21
Y20 <- AA$Y[20]
e20 <- Y20 - (b0 + b1 * X20)
F21 <- Yhat21 + rho*e20

alpha <- 0.01
n <- length(AA$Y)
s.pred <- sqrt(MSE*(1 + (1/n) + (X21p - Xbar.p)^2/(sum((Xp[-1] - Xbar.p)^2))))
s.pred
pred.L <- F21 - qt(1-alpha/2, df=n-3)*s.pred
pred.U <- F21 + qt(1-alpha/2, df=n-3)*s.pred

print(paste(100*(1-alpha) ,"percent prediction interval for response at month 21 is:", round(pred.L,4), ",", round(pred.U,4) ))
```
Given the staff time in month 21 as 3.625 thousand hours, and given the amount of billings in month 20
as 278.5 thousand constant dollars, we predict with approximately 99 percent confidence that
the amount of billings in month 21 will be between 276.98 and 279.73 thousand constant dollars

## Problem 3

a-)
```{r P3_a}
hildreth <- function(x,y,rho){
  n3 <- length(rho)
  out <- matrix(0, nrow=n3, ncol=2)
  out[,1] <- rho

  for (i in 1:n3){
    d <- anova(hildreth.lu(y=y, x=x, rho=rho[i]))
    out[i,2] <- d$"Sum Sq"[2]
  }
  out
}
rh <- seq(0.1,1,by=0.01)
hl <- hildreth(AA$X, AA$Y, rh)
plot(hl, xlab=expression(rho), ylab="SSE")
rho3a <- hl[which.min(hl[,2]),][1] 

Ytp <- AA$Y - rho3a * Lag(AA$Y, shift = 1)
Xtp <- AA$X - rho3a * Lag(AA$X, shift = 1)

f3.1 <- lm(Ytp ~ Xtp)
summary(f3.1)
b1.3a <- f3.1$coefficients[2]
sb1.3a <- summary(f3.1)$coefficients[2,2]
```

The value of $\rho$ that minimized SSE is `r round(rho3a,2)`.
The estimated transformed regression function:\
$\hat{Y'}$ = 54.2435 + 50.4659X'

b-)
```{r P3_b}
rho3b <- 1
Ytp2 <- AA$Y - rho3b*Lag(AA$Y, shift = 1)
Xtp2 <- AA$X - rho3b*Lag(AA$X, shift = 1)
f3.2 <- lm(Ytp2 ~ Xtp2 - 1)
summary(f3.2)
b1.3b <- f3.2$coefficients[1]
b0.3b <- mean(AA$Y) - mean(AA$X)*b1.3b
sb1.3b <- summary(f3.2)$coefficients[1,2]
```

The first difference procedure uses r = 1, and the estimated transformed regression function is:\
$\hat{Y'}$ = 50.164X'

c-)
The Durbin-Watson test results below show that the autocorrelation is successfully remediated
for both models.
```{r P3_c}
dwtest(f3.1) # Hildreth-Lu
dwtest(f3.2) # First Difference
```

$H_o$: $\rho$ = 0 \
$H_a$: $\rho$ > 0 

For both cases, the p-value is greater than $\alpha$=0.01, so we fail to reject 
the null hypothesis. There seems to be no autocorrelation.

d-)
```{r P3_d}
MSE1 <- summary(f3.1)$sigma^2 # Hildreth-Lu
MSE2 <- summary(f3.2)$sigma^2 # First difference

hild <- c(b1.3a, sb1.3a, rho3a, MSE1)
fd <- c(b1.3b, sb1.3b, rho3b, MSE2)
tbl <- data.frame(rbind(hild, fd))
rownames(tbl) <- c("Hildreth-Lu", "First difference")
colnames(tbl) <- c("b1", "s{b1}", "r", "MSE")

kable(tbl, digits=4, align="c", 
      booktabs=TRUE) %>% kable_styling(latex_options="HOLD_position")
```

Based on the table above, we would choose the Hildreth-Lu since it resulted in 
smaller MSE and $s{b_1}$.

e-)
```{r P3_e}
Xp <- Xtp
Xbar.p <- mean(Xp[-1])

X20 <- AA$X[20]
X21 <- 3.625
X21p <- X21 - rho3a * X20

Yhat21 <- b0 + b1 * X21
Y20 <- AA$Y[20]
e20 <- Y20 - (b0 + b1 * X20)
F21 <- Yhat21 + rho3a * e20

alpha <- 0.01
n <- length(AA$Y)
s.pred <- sqrt(MSE1*(1 + (1/n) + (X21p - Xbar.p)^2/(sum((Xp[-1] - Xbar.p)^2))))
s.pred
pred.L <- F21 - qt(1-alpha/2, df=n-3)*s.pred
pred.U <- F21 + qt(1-alpha/2, df=n-3)*s.pred

print(paste(100*(1-alpha) ,"percent prediction interval for response at month 21 is:", round(pred.L,4), ",", round(pred.U,4)))
```
Given the staff time in month 21 as 3.625 thousand hours, and given the amount of billings in month 20
as 278.5 thousand constant dollars, we predict with approximately 99 percent confidence that
the amount of billings in month 21 will be between 277.07 and 279.79 thousand constant dollars.


## Problem 4

a-)
There are many zeros in the dataset that do not actually make sense. We consider
these as recording errors and remove them.

```{r P4_a}
data(pima)
str(pima)
head(pima)

# Set 0's to NA since 0 values do not make sense
pima$diastolic[pima$diastolic == 0] <- NA
pima$glucose[pima$glucose == 0] <- NA
pima$triceps[pima$triceps == 0] <- NA
pima$insulin[pima$insulin == 0] <- NA
pima$bmi[pima$bmi == 0] <- NA
pima$test <- as.factor(pima$test)
apply(pima,2,function(x) sum(is.na(x)))
pima2 <- na.omit(pima[,-5])

set.seed(456)
train.index4 <- createDataPartition(y = pima2$test, p = 0.7, list = FALSE)
train.set4 <- pima2[train.index4,]
test.set4 <- pima2[-train.index4,]

f.tree4 <- rpart(test ~ ., data=train.set4, method="class")
rpart.plot(f.tree4, digits=3)
f.tree4$cptable
plotcp(f.tree4)

a4.1 <- defaultSummary(data.frame(obs = pima2$test, pred = predict(f.tree4, pima2, type="class")))

f.pruned4 <- prune(f.tree4, cp = getBestCp(f.tree4))
rpart.plot(f.pruned4, digits=3)
f.pruned4$cptable
plotcp(f.pruned4)

a4.2 <- defaultSummary(data.frame(obs = pima2$test, pred = predict(f.pruned4, pima2, type="class")))
table4 <- rbind(a4.1, a4.2)
rownames(table4) <- c("Regression Tree","Pruned Regression Tree")
kable(table4, digits=3, align="c", booktabs=TRUE,caption="Model Comparison") %>%
  kable_styling(latex_options="HOLD_position")
```
The $C_p$ table for the regression tree suggests that we prune the tree, but the
pruned tree happens to be overly simple and has only one split. The model comparison
table shows that the normal regression tree actually performs better with higher 
accuary and kaap when evaluated on the full data. 

b-)
```{r P4_b}
set.seed(456)
bg <- ipred::bagging(test ~ ., data = train.set4, nbagg = 25)
pred.bg <- predict(bg, pima2)
b4.1 <- defaultSummary(data.frame(obs = pima2$test, pred = pred.bg))

ctrl <- trainControl(method = "cv", number = 10)
set.seed(456)
bgt <- train(form = test ~ .,
                         data = train.set4,
                         method = "treebag",
                         preProcess = c("center", "scale"),
                         metric="Accuracy",
                         tuneLength=20,
                         trControl=ctrl)
pred.bgt <- predict(bgt, pima2)
b4.2 <- defaultSummary(data.frame(obs = pima2$test, pred = pred.bgt))

table5 <- rbind(a4.1, a4.2, b4.1, b4.2)
rownames(table5) <- c("Regression Tree","Pruned Regression Tree","Bagging 1","Bagging 2")
kable(table5, digits=3, align="c", booktabs=TRUE,caption="Model Comparison") %>%
  kable_styling(latex_options="HOLD_position")
```

We took two different approaches for bagging, both of which performed well with 
similar accuracy and kappa. The bagging method looks better than the regression tree
in predicting the diabetes test results.

c-)
```{r P4_c}
set.seed(456)
ctrl <- trainControl(method = "cv", number = 10)
bst <- train(form = test ~ .,
            data = train.set4,
            method = "blackboost",
            preProcess = c("center", "scale"),
            metric="Accuracy",
            trControl=ctrl)
pred.bst <- predict(bst, pima2)
c4 <- defaultSummary(data.frame(obs = pima2$test, pred = pred.bst))

table6 <- rbind(a4.1, a4.2, b4.1, b4.2, c4)
rownames(table6) <- c("Regression Tree","Pruned Regression Tree","Bagging 1","Bagging 2","Boosting")
kable(table6, digits=3, align="c", booktabs=TRUE,caption="Model Comparison") %>%
  kable_styling(latex_options="HOLD_position")
```
The boosting method performed worse than the bagging, so we still keep the bagging
model as the best one at this point.

d-)
```{r P4_d}
rf <- randomForest(test ~ ., data = train.set4)
pred.rf <- predict(rf, pima2)
d4 <- defaultSummary(data.frame(obs = pima2$test, pred = pred.rf))

table7 <- rbind(a4.1, a4.2, b4.1, b4.2, c4, d4)
rownames(table7) <- c("Regression Tree","Pruned Regression Tree","Bagging 1","Bagging 2","Boosting","Random Forest")
kable(table7, digits=3, align="c", booktabs=TRUE,caption="Model Comparison") %>%
  kable_styling(latex_options="HOLD_position")
```

The final table shows that the random forest model is also very good at predicting, 
and the accuracy and kappa are very close to the two bagging models. Strictly based
on the numbers shown in the table, the fist bagging model shows the best results, 
but because the difference is very marginal, we will consider both the bagging and
random forest as the final candidate and do further analysis.